### 2010-12-04 paper.tex
Added latex paper

% !TEX TS-program = pdflatex
% !TEX encoding = UTF-8 Unicode

% This is a simple template for a LaTeX document using the "article" class.
% See "book", "report", "letter" for other types of document.

\documentclass[11pt]{article} % use larger type; default would be 10pt

\usepackage[utf8]{inputenc} % set input encoding (not needed with XeLaTeX)

%%% Examples of Article customizations
% These packages are optional, depending whether you want the features they provide.
% See the LaTeX Companion or other references for full information.

%%% PAGE DIMENSIONS
\usepackage{geometry} % to change the page dimensions
\geometry{a4paper} % or letterpaper (US) or a5paper or....
% \geometry{margins=2in} % for example, change the margins to 2 inches all round
% \geometry{landscape} % set up the page for landscape
%   read geometry.pdf for detailed page layout information

\usepackage{graphicx} % support the \includegraphics command and options

% \usepackage[parfill]{parskip} % Activate to begin paragraphs with an empty line rather than an indent

%%% PACKAGES
\usepackage{booktabs} % for much better looking tables
\usepackage{array} % for better arrays (eg matrices) in maths
\usepackage{paralist} % very flexible & customisable lists (eg. enumerate/itemize, etc.)
\usepackage{verbatim} % adds environment for commenting out blocks of text & for better verbatim
\usepackage{subfig} % make it possible to include more than one captioned figure/table in a single float
\usepackage{amstext}
% These packages are all incorporated in the memoir class to one degree or another...

%%% HEADERS & FOOTERS
\usepackage{fancyhdr} % This should be set AFTER setting up the page geometry
\pagestyle{fancy} % options: empty , plain , fancy
\renewcommand{\headrulewidth}{0pt} % customise the layout...
\lhead{}\chead{}\rhead{}
\lfoot{}\cfoot{\thepage}\rfoot{}

%%% SECTION TITLE APPEARANCE
\usepackage{sectsty}
\allsectionsfont{\sffamily\mdseries\upshape} % (See the fntguide.pdf for font help)
% (This matches ConTeXt defaults)

%%% ToC (table of contents) APPEARANCE
\usepackage[nottoc,notlof,notlot]{tocbibind} % Put the bibliography in the ToC
\usepackage[titles,subfigure]{tocloft} % Alter the style of the Table of Contents
\renewcommand{\cftsecfont}{\rmfamily\mdseries\upshape}
\renewcommand{\cftsecpagefont}{\rmfamily\mdseries\upshape} % No bold!

%%% END Article customizations

%%% The "real" document content comes below...

\title{Minimizing Time Required For Graduation}
\author{Ruyi Cai \and Matthew Fonda \and Michael Hughes}
%\date{} % Activate to display a given date or no date (if empty),
         % otherwise the current date is printed

\begin{document}
\maketitle

% Adds a table of contents. Comment out if we don't want this.
\tableofcontents
\pagebreak

\abstract{The Bachelor's Degree is designed to take four years to complete; however, many students fail to complete it this amount of time. One of the factors contributing to this is a lack of proper planning during the registration phase. In this paper we present an algorithm to assist students in constructing a schedule that works towards completing a degree in the shortest amount of time possible.}

\section{Introduction}
\subsection{Background}
Registration can be a stressful time for every student at the Univeristy of Washington. Without deliberate planning, students must decide which classes to take on a quarterly basis. The decision process involves considering many factors, such as quarterly class offerings, the amount of credits, the prerequisites of the classes, time conflicts, and how the choices in current quarter affect future quarters. The University of Washington registration system provides a schedule finder; however, it only takes into account time conflicts, and does not give advice on how to pick the most ideal schedule. The situation described above makes it difficult for students advising faculties. A student's graduation time may be extended because of his or her inability to take a class that has limited offerings during that particular year, for reasons such as not satisfying the prerequisites of the class. The advising office’s workload can also be increased due to the changing plans and majors of students and their need for repeated advising.

\subsection{Goal}
Our goal is to alleviate the problematic situation by creating a mathematical algorithm that accomplishes the following goal: given a student's current state and preferred credit hours per quarter, produce a class schedule that minimizes the number of quarters until graduation, eliminates time conflicts, and eliminates prerequiste conflicts.

\subsection{Importance}
The implementation of the algorithm as part of the school’s registration system will benefit the students by enabling them to map out their quarterly class schedules until graduation, and to preview their ideal schedule, the length of time until graduation, and the credit hours per quarter. This will lead to better decision making for what majors to pursue, and given the student's current state, which schedules are the most ideal in terms of saving time and resources. This will also help studentssave money, as they will be in school for fewer quarters. The university also benefits from the reduced labor cost and increased efficiency of the advising offices.

\section{Problem Description}
We focus on the classes offered within the Math department, although our algorithm and model can be easily extended to include over departments.

Every quarter the Math department offers a subset of classes from all of the available courses, each class is offered during a certain subset of available time periods. Since there are far more classes than time periods during the regular five-day week there are inevitably conflicts between courses that a student wants to take and courses that a student needs to take. Or worse, there are conflicts between two or more courses that a students needs to graduate. To graduate with a B.S. in Mathematics there are certain classes that are required and then there are elective requirements too. Additionally a student may wish to take some classes that are not required.
Adding to the problem is many classes cannot just be taken whenever they are available; instead, certain prerequisites must be satisfied first. In the case of math courses, there can be multiple different prerequisite sets that satisfy the requirements for a class.

Our goal is to determine a schedule of classes from some starting point to graduation that given the above properties minimizes the total number of quarters needed. Slightly more formally we would like to: \\
Minimize the total number of quarters needed to graduate subject to:
\begin{enumerate}
\item Only taking courses during quarters they are available
\item Only taking a course once
\item Never exceeding a preset number of credit hours per quarter
\item Never taking more than one class at a given time
\item Always satisfying a course's prerequisites prior to taking that course
\end{enumerate}

\subsection{Definitions}
We will now formalize some terms used in our paper.
\begin{itemize}
\item A {\it course} refers to any course offered by the University of Washington. A course has associated with it a name used in the University of Washington time schedule, for example {\it MATH309}. A course also has associated with it the quarters in which it is offered. Throughout this paper, we will refer to these using their time schedule abbreviates: {\it F} for Fall Quarter, {\it W} for Winter Quarter, {\it Sp} for Spring Quarter, and {\it S} for Summer Quarter. Each course is also worth a specific amount of credits.
\item We refer to a specific instance of a course as a {\it course offering}. Each course offering additionally has associated with it the time of the day at which the course meets, the quarter it is being offered, and the section number of the offering.
\end{itemize}

We must now also define some variables used throughout our paper.
\begin{itemize}
\item Let $S$ be the set of courses required to be taken in order to graduate. $S$ can be thought of as our goal state, and can also contain elective courses a student simply wishes to take.
\item Let $T$ be the set of courses a given student has already completed. If no courses have been taken, this set may be empty.
\item Let $U$ be the set of all courses a student could possibly ever take.
\item Let $P$ be a set of rules specifying course prerequistes.
\item Let $l$ be the maximum number of credits a student wishes to take during any given quarter.
\item Let $O$ be the set of all 50-minute time blocks between 8:30am and 7:30pm from Monday through Friday.
\end{itemize}

\subsection{Representing Prerequisites}
Initially, we hoped to represent course prerequistes using a directed acyclic graph, in which each node was a course, and edges between nodes $i$ and $j$ if and only if $i$ was a prerequisite for $j$.  This method was used by Pritsker et. al. and Xiangtong et. al. However, it does not work in our model, and courses can have a complicated set of multiple prerequisites. For example, {\it MATH327} requires both $MATH126$ and $MATH300$, or just $MATH136$. Let this graph be called $G_{prereq}$. Due to these complications, a $G_{prereq}$ alone will not suffice.

To solve this issues, we introduce a set of rules for each course. Each rule is a boolean expression represented in propositional logic in disjunctive normal form. For example, the rule assoicated with $MATH327$ is
$$
\text{{\it MATH327: (MATH126 AND MATH300) OR MATH136}}.
$$
As previously stated, we denote the set of all rules for all classes as $P$.

Furthermore, to determine if a given student is eligible to register for a course offering, we can evaluate the rule for the course, where each clause takes on the value true if a student has already taken the course, and false if he or she hasn't. If the rule evaluates to true, then the student is eligible to take the course; otherwise, the student is not.

\subsection{Initial Model and Tracability}
With a static schedule it would be possible to deterministically plan an entire schedule from start to finish, unfortunately we have neither the computational resources nor the luxury of a static schedule. From the student’s point of view, although there are some classes that are offered every quarter, senior electives are not, and sometimes it cannot be predicted more than one quarter in advance when a senior elective will be offered. Even given perfect information, where we know the possible class schedules for every quarter we face another problem; finding the best possible schedule reduces to enumerating possible class schedules for every quarter until the graduation requirements are satisfied. Given that we cannot predict in advance the conflicts between classes, we cannot even determine the branching factor of the enumeration tree nor its depth. Since we can only determine one quarter in advance, this suggests approaching the problem by using heuristics to pick the best possible classes each quarter.

Instead of trying to solve the complete problem we chose to use the rolling horizon approach used by Xiangtong et al \cite{xiangton:informs} when scheduling classes for airlines. Our problem already has a well defined horizon demarcation, the quarters. For the rest of the paper we focus on developing methods for attempting to optimally pick classes in each quarter separately.

\section{Rolling Horizons Approach}
We adapt a rolling horizons approach which uses a heuristic to pick the best possible schedule for any given quarter, given the classes offered during the quarter and the courses the student has previously taken. We use a greedy algorithm that functions on a per-quarter basis to accomplish this. The greedy algorithm greatly improves the tracability of course scheduling over an initially unknown number of quarters.

We break our algorithm down into two main phases: determining available course offering and selecting specific course offerings.

\subsection{Determining Available Course Offerings}
We need to select the classes we can currently take and then build conflict graphs for classes based on the times they are offered. In this section, we let $L$ be the initially empty set of course offerings a given student is eligible to take in the given quarter, $q$. \\

\subsubsection{Procedue}
We present the procedure for popluating the set $L$ as follows:
\begin{enumerate}
\item For each rule in $P$ we iterate over each proposition in the rule, $p$, and check if $p \in T$. If $p \in T$ then we say $p$ evaluates to true; otherwise, $p$ evaluates to false. We recall that each rule is associated with a given course. If the entire rule evaluates to true, then we add the course associated with it to $L$. Once this is complete, $L$ will consist of every course that could be taken based solely on prerequisites that we have fulfilled.
\item We now prune this set by removing all courses that are not offered in the current quarter. $L$ will necessarily now only contain classes that we can take given the current quarter and the classes we have already taken.
\item We now construct a confict graph, $G_{conflict}$, where every course offering in $L$ is a node, and there is an edge between two nodes $i$ and $j$ if and only if the time of course offering $i$ overlaps the time of course offering $j$.
\end{enumerate}

\subsubsection{Analysis}
The selection procedure involves a lot of set queries and iteration, but in asymptotic running time the algorithm is relatively quick. Let $n$ be the total number of classes ever available to take, or in other words, $n = \|U\|$. In the worst case, $P$ has a complex rule for all $n$ courses, and checking each rule in $P$ is linear in the number of propositions. However, we know {\it a priori} from analysis of our dataset that the largest rule has twelve propositions, so we can bound the rule check time by a constant, $C$, and therefore step one can be done in $O(Cn)$ time.

Continuing to step two, the worst case is when $L$ = $Q$. Checking whether a course is offered in a given quarter can be done in constant time, and we have $n$ courses to check, therefore step two can also be done in $O(Cn)$ time.

Finally, the last step is once again bounded by a constant and $n$. Considering courses between 8:30am and 7:30pm, we have nine fifty-minute time periods, and five days of the week, and therefore we have 45 time slots to check in our conflict graph, again yielding $O(Cn)$ runtime.

Since we are dealing with asymptotic running time we will drop the constant terms and combining all resulsts yields an overall runtime of $O(n)$.

\subsection{Selection of Quarterly Classes}
\subsubsection{Selection Prodcedure}
At this point in time, we know which course offerings are available for the given student to register for, and we must now pick the most optimal ones. There a several ways we could view the problem of determining which classes to take each quarter. We will initially encode the class selection as a binary integer programming problem in spirit of the methods presented by Dinkel et al \cite{dinel:scheduling}. and Pritsker et al \cite{prisker:informs}. At this point we have the set $L$ containing possible classes we that we need to take.

\subsubsection{Heuristic Value Function}
In order to determine which classes should be picked in a given schedule, we use a heuristic function which assigns each course 
offering a value. The higher the value of this function for a given course offering is, the more likely the given course offering 
will be placed in the schedule. Recalling that our goal is to minimize the total number of quarters taken (and implicitly, 
maximize the total number of courses taken per quarter), our heuristic function should clearly give higher value to course 
offerings that will help us achieve this goal. One component to the heuristic function then must be the number of dependent 
courses a given course offering has. If a course satisfies many prerequisties, then we know that there are many courses that 
we cannot complete without it, so we want our heuristic function to favor it.

Additionally, we know that not every course is offered every quarter. If a course is not offered until the following year, we 
could potentially add an entire year to the schedule. This is a situation we hope to avoid. Our heuristic function must 
therefore consider how often and how far in the future a given course is offered. For example, if we are deciding between 
two course offering that are identical except one is ofference once a year, and the other is offered three times a year, we 
want to take the course only offered once a year, as we won't have the oppurtunity to take it for another entire year.

We these considerations in mind, we begin to develop our heuristic function. In order to define our heuristic function, we 
must first define some auxilary functions on course offering.
\begin{itemize}
\item Let $TimeFromNow(c, q)$ be difference in quarters between the current quarter, $q$, and the next quarter a given course offering, $c$, is offered. As there are four quarters, the range of this function is $\{0, 1, 2, 3\}$.
\item Let $Quarters(c)$ be the number of quarters per year a given course, $c$, is offered. As there are four quarters, the range of this function is $\{0, 1, 2, 3, 4\}$.
\item Let $Dependents(c)$ be the number of course that either directly or indirectly require a given course, $c$. In our final 
draft, we will present a diagram illustrating this value more clearly.
\end{itemize}

\includegraphics{more_prereq_tree}

With these functions defined, we can now define our heuristic function:
\begin{equation}
Value(c) = 1 + \alpha * Dependents(c) + \beta * TimeFromNow(c, q) + \gamma * Quarters(c)
\end{equation}
Where $c$ is a given course offering, $q$ is the current quarter, and $\alpha$, $\beta$, and $\gamma$ are real numbers, giving a 
weight to piece of the heuristic. The scalars are determined experimentally.

\subsubsection{Formulation as Binary Integer Program}
We can now formulate our problem as a binary integer linear programming probelm. Before stating the linear program, we must first define some variables used in it: \\
\begin{itemize}
\item $
  X_i = \left\{
     \begin{array}{lr}
       1 \text{ if we take } X_i\\
       0 \text{ if we don't take } X_i
     \end{array}
   \right. $
\item For all $t \in O$, we let $t = 1$.
\item Let $m = \|L\|$.
\end{itemize}
We now state our problem as follows:\\
Maximize:
$$
\text{Value of Quarter} = \sum_{i=1}^m Value(c_i) * X_i
$$
Subject To:
$$
\sum_{i=1}^m Weight(c_i) * X_i \leq l,
$$
$$
\forall t_j \in O, \sum_{i=1}^m t_j * X_i \leq 1,
$$
$$
X_i \in \{0, 1\}
$$
The linear program can be interpreted as follows:
\begin{itemize}
\item The value of a quarter is the total number of courses taken that quarter.
\item In each quarter, the total number of credits cannot exceed the specified limit.
\item Only one course can be taken in one particular time slot
\item A course can only be taken or not taken. In other words, decision variables are binary.
\item Each non-zero descision variable in the optimal solution represents a class we will take.
\end{itemize}

It is worth nothing that all of the t variables are essentially dummy variables that ensure the correct number of equations of type 2. Each of the equations based on eqn. 2 ensure that we only take one class per time period per day. Also this linear program is ‘nice’ in that it always has feasible origin (don’t take any classes), and that the simplest formulation is already in standard form.

As noted in the reference literature this is actually a (strongly) NP-complete problem as described \cite{pferschy:kcg}. For the rest of this section we consider an alternative formulation of class selection as a knapsack problem with conflict graphs and suggest the use a technique developed by Yamada et. al. \cite{yamada:heuristic}.

We begin by simplifying $G_{conflict}$ to be a fully connected graph by adding dummy nodes $v'$ and insert edges that connect $v'$ to every other node in the graph. To ensure that $v'$ does not alter the final optimal solution, we assign it the following properties:
\begin{itemize}
\item $Weight(v') = l+1$
\item $Value(v') = 0$
\end{itemize}
As the weight of our dummy node is greater than our credit limit, and it's value is zero, we know that it will never appear in the optimal solution. With the above reasoning we show for our case of class schedule that Pferschy and Schauer's \cite{pferschy:kcg} claim that adding a dummy node without loss of generality holds. We can now conveniently write the class picking problem as the knapsack with conflict graphs problem (KCG): \\
Maximize:
$$
\text{Value of Quarter} = \sum_{i=1}^m Value(c_i) * X_i
$$
Subject To:
$$
\sum_{i=1}^m Weight(c_i) * X_i \leq l,
$$
$$
X_i + X_j \leq 1 \text{ for all connected vertices } i, j \in G_{conflict},
$$
$$
X_i \in \{0, 1\}
$$
It is worth noting that for relatively small versions of this problem, in the low hundreds of variables it is possible to solve this problem exactly using modern LP solving software \cite{yamada:heuristic}. In our case, just solving for the optimal classes for a single major fits well into the scope of a few hundred (or even just tens of) variables. For multiple majors or to consider class selection for distribution and elective requirements, however the number of classes balloons significantly and the KCG problem becomes much harder. For this version of the problem we suggest using the techniques used by Yamada et al \cite{yamada:heuristic}, which we briefly outline below.

\subsubsection{Relaxation and Implicit Enumeration}
\begin{enumerate}
\item Develop upper bound on value by solving a continuous version knapsack problem generated by a Lagrangean relaxation of our problem. Let $X_i$ be a continuous variable with $0 \leq X_i \leq 1$, and shift the time constraints into the objective function by using a penalty of $\lambda$ when the constraint is violated.
\item Use the upper bound developed above with the pruning rules suggested by Yamada et al \cite{yamada:heuristic} to do implicit enumeration of the possible choices.
\end{enumerate}

In Yamada et al’s \cite{yamada:heuristic} numerical runtime analysis they found this algorithm ran very favorably against solving the same problem using popular linear programming solvers. There is one problem that we should briefly address, because the number of periods of available classes per week is relatively small the conflict rate for a large number of class is relatively higher than Yamada et al’s 2\% maximal conflict rate for their experiments. Our data has a significant mitigating factor though, knapsack problems in general are exponential in terms of the number of choices and the total weight capacity. Realistically the total weight capacity is going to be 20 credits or less which is between 1\% and 2\% of the figures used by Yamada \cite{yamada:heuristic} and also our total number of choices is in many cases (particular just the classes in a particular major) less than 200. Giving the size and type of data in our set the higher conflict rate is less of a concern and Yamada et al’s algorithm should still be efficient.

\section{Computational Results}
In this section we will outline results obtained by running computer simulations of schedule picking. We will computationally examine what parameters in our heuristic function produce the most desirable schedules.

\bibliography{references}{}
\bibliographystyle{plain}
\end{document}

### 2010-12-04 paper.tex
made paper.tex readable by limiting line length to 80 chars

@@
 
-\abstract{The Bachelor's Degree is designed to take four years to complete; however, many students fail to complete it this amount of time.
-One of the factors contributing to this is a lack of proper planning during the registration phase.
-In this paper we present an algorithm to assist students in constructing a schedule that works towards completing a degree in the shortest amount of time possible.}
+\abstract{The Bachelor's Degree is designed to take four years to complete;
+however, many students fail to complete it this amount of time. One of the
+factors contributing to this is a lack of proper planning during the
+registration phase. In this paper we present an algorithm to assist students in
+constructing a schedule that works towards completing a degree in the shortest
+amount of time possible.}
 
-\section{Introduction}
-\subsection{Background}
-Registration can be a stressful time for every student at the Univeristy of Washington.
-Without deliberate planning, students must decide which classes to take on a quarterly basis.
-The decision process involves considering many factors, such as quarterly class offerings, the amount of credits, the prerequisites of the classes, time conflicts, and how the choices in current quarter affect future quarters.
-The University of Washington registration system provides a schedule finder; however, it only takes into account time conflicts, and does not give advice on how to pick the most ideal schedule.
-The situation described above makes it difficult for students advising faculties.
-A student's graduation time may be extended because of his or her inability to take a class that has limited offerings during that particular year, for reasons such as not satisfying the prerequisites of the class.
-The advising office’s workload can also be increased due to the changing plans and majors of students and their need for repeated advising.
+\section{Introduction} \subsection{Background} Registration can be a stressful
+time for every student at the Univeristy of Washington. Without deliberate
+planning, students must decide which classes to take on a quarterly basis. The
+decision process involves considering many factors, such as quarterly class
+offerings, the amount of credits, the prerequisites of the classes, time
+conflicts, and how the choices in current quarter affect future quarters. The
+University of Washington registration system provides a schedule finder;
+however, it only takes into account time conflicts, and does not give advice on
+how to pick the most ideal schedule. The situation described above makes it
+difficult for students advising faculties. A student's graduation time may be
+extended because of his or her inability to take a class that has limited
+offerings during that particular year, for reasons such as not satisfying the
+prerequisites of the class. The advising office’s workload can also be increased
+due to the changing plans and majors of students and their need for repeated
+advising.
 
-\subsection{Goal}
-Our goal is to alleviate the problematic situation by creating a mathematical algorithm that accomplishes the following goal: given a student's current state and preferred credit hours per quarter, produce a class schedule that minimizes the number of quarters until graduation, eliminates time conflicts, and eliminates prerequiste conflicts.
+\subsection{Goal} Our goal is to alleviate the problematic situation by creating
+a mathematical algorithm that accomplishes the following goal: given a student's
+current state and preferred credit hours per quarter, produce a class schedule
+that minimizes the number of quarters until graduation, eliminates time
+conflicts, and eliminates prerequiste conflicts.
 
-\subsection{Importance}
-The implementation of the algorithm as part of the school’s registration system will benefit the students by enabling them to map out their quarterly class schedules until graduation, and to preview their ideal schedule, the length of time until graduation, and the credit hours per quarter.
-This will lead to better decision making for what majors to pursue, and given the student's current state, which schedules are the most ideal in terms of saving time and resources.
-This will also help studentssave money, as they will be in school for fewer quarters.
-The university also benefits from the reduced labor cost and increased efficiency of the advising offices.
+\subsection{Importance} The implementation of the algorithm as part of the
+school’s registration system will benefit the students by enabling them to map
+out their quarterly class schedules until graduation, and to preview their ideal
+schedule, the length of time until graduation, and the credit hours per quarter.
+This will lead to better decision making for what majors to pursue, and given
+the student's current state, which schedules are the most ideal in terms of
+saving time and resources. This will also help studentssave money, as they will
+be in school for fewer quarters. The university also benefits from the reduced
+labor cost and increased efficiency of the advising offices.
 
-\section{Problem Description}
-We focus on the classes offered within the Math department, although our algorithm and model can be easily extended to include over departments.
+\section{Problem Description} We focus on the classes offered within the Math
+department, although our algorithm and model can be easily extended to include
+over departments.
 
-Every quarter the Math department offers a subset of classes from all of the available courses, each class is offered during a certain subset of available time periods.
-Since there are far more classes than time periods during the regular five-day week there are inevitably conflicts between courses that a student wants to take and courses that a student needs to take.
-Or worse, there are conflicts between two or more courses that a students needs to graduate.
-To graduate with a B.S.
-in Mathematics there are certain classes that are required and then there are elective requirements too.
-Additionally a student may wish to take some classes that are not required.
-Adding to the problem is many classes cannot just be taken whenever they are available; instead, certain prerequisites must be satisfied first.
-In the case of math courses, there can be multiple different prerequisite sets that satisfy the requirements for a class.
+Every quarter the Math department offers a subset of classes from all of the
+available courses, each class is offered during a certain subset of available
+time periods. Since there are far more classes than time periods during the
+regular five-day week there are inevitably conflicts between courses that
+a student wants to take and courses that a student needs to take. Or worse,
+there are conflicts between two or more courses that a students needs to
+graduate. To graduate with a B.S. in Mathematics there are certain classes that
+are required and then there are elective requirements too. Additionally
+a student may wish to take some classes that are not required.  Adding to the
+problem is many classes cannot just be taken whenever they are available;
+instead, certain prerequisites must be satisfied first. In the case of math
+courses, there can be multiple different prerequisite sets that satisfy the
+requirements for a class.
 
-Our goal is to determine a schedule of classes from some starting point to graduation that given the above properties minimizes the total number of quarters needed.
-Slightly more formally we would like to: \\
-Minimize the total number of quarters needed to graduate subject to:
-\begin{enumerate}
-\item Only taking courses during quarters they are available
-\item Only taking a course once
-\item Never exceeding a preset number of credit hours per quarter
-\item Never taking more than one class at a given time
-\item Always satisfying a course's prerequisites prior to taking that course
-\end{enumerate}
+Our goal is to determine a schedule of classes from some starting point to
+graduation that given the above properties minimizes the total number of
+quarters needed. Slightly more formally we would like to: \\ Minimize the total
+number of quarters needed to graduate subject to: \begin{enumerate} \item Only
+taking courses during quarters they are available \item Only taking a course
+once \item Never exceeding a preset number of credit hours per quarter \item
+Never taking more than one class at a given time \item Always satisfying
+a course's prerequisites prior to taking that course \end{enumerate}
 
-\subsection{Definitions}
-We will now formalize some terms used in our paper.
-\begin{itemize}
-\item A {\it course} refers to any course offered by the University of Washington.
-A course has associated with it a name used in the University of Washington time schedule, for example {\it MATH309}.
-A course also has associated with it the quarters in which it is offered.
-Throughout this paper, we will refer to these using their time schedule abbreviates: {\it F} for Fall Quarter, {\it W} for Winter Quarter, {\it Sp} for Spring Quarter, and {\it S} for Summer Quarter.
-Each course is also worth a specific amount of credits.
+\subsection{Definitions} We will now formalize some terms used in our paper.
+\begin{itemize} \item A {\it course} refers to any course offered by the
+University of Washington. A course has associated with it a name used in the
+University of Washington time schedule, for example {\it MATH309}. A course also
+has associated with it the quarters in which it is offered. Throughout this
+paper, we will refer to these using their time schedule abbreviates: {\it F} for
+Fall Quarter, {\it W} for Winter Quarter, {\it Sp} for Spring Quarter, and {\it
+S} for Summer Quarter. Each course is also worth a specific amount of credits.
 \item We refer to a specific instance of a course as a {\it course offering}.
-Each course offering additionally has associated with it the time of the day at which the course meets, the quarter it is being offered, and the section number of the offering.
-\end{itemize}
+Each course offering additionally has associated with it the time of the day at
+which the course meets, the quarter it is being offered, and the section number
+of the offering.  \end{itemize}
 
 We must now also define some variables used throughout our paper.
-\begin{itemize}
-\item Let $S$ be the set of courses required to be taken in order to graduate. $S$ can be thought of as our goal state, and can also contain elective courses a student simply wishes to take.
-\item Let $T$ be the set of courses a given student has already completed. If no courses have been taken, this set may be empty.
-\item Let $U$ be the set of all courses a student could possibly ever take.
-\item Let $P$ be a set of rules specifying course prerequistes.
-\item Let $l$ be the maximum number of credits a student wishes to take during any given quarter.
-\item Let $O$ be the set of all 50-minute time blocks between 8:30am and 7:30pm from Monday through Friday.
-\end{itemize}
+\begin{itemize} \item Let $S$ be the set of courses required to be taken in
+order to graduate. $S$ can be thought of as our goal state, and can also contain
+elective courses a student simply wishes to take.  \item Let $T$ be the set of
+courses a given student has already completed. If no courses have been taken,
+this set may be empty.  \item Let $U$ be the set of all courses a student could
+possibly ever take.  \item Let $P$ be a set of rules specifying course
+prerequistes.  \item Let $l$ be the maximum number of credits a student wishes
+to take during any given quarter.  \item Let $O$ be the set of all 50-minute
+time blocks between 8:30am and 7:30pm from Monday through Friday.  \end{itemize}
 
-\subsection{Representing Prerequisites}
-Initially, we hoped to represent course prerequistes using a directed acyclic graph, in which each node was a course, and edges between nodes $i$ and $j$ if and only if $i$ was a prerequisite for $j$.
-This method was used by Pritsker et.
-al.
-and Xiangtong et.
-al.
-However, it does not work in our model, and courses can have a complicated set of multiple prerequisites.
-For example, {\it MATH327} requires both $MATH126$ and $MATH300$, or just $MATH136$.
-Let this graph be called $G_{prereq}$.
-Due to these complications, a $G_{prereq}$ alone will not suffice.
+\subsection{Representing Prerequisites} Initially, we hoped to represent course
+prerequistes using a directed acyclic graph, in which each node was a course,
+and edges between nodes $i$ and $j$ if and only if $i$ was a prerequisite for
+$j$.  This method was used by Pritsker et. al. and Xiangtong et. al. However, it
+does not work in our model, and courses can have a complicated set of multiple
+prerequisites. For example, {\it MATH327} requires both $MATH126$ and $MATH300$,
+or just $MATH136$. Let this graph be called $G_{prereq}$. Due to these
+complications, a $G_{prereq}$ alone will not suffice.
 
-To solve this issues, we introduce a set of rules for each course.
-Each rule is a boolean expression represented in propositional logic in disjunctive normal form.
-For example, the rule assoicated with $MATH327$ is
-$$
-\text{{\it MATH327: (MATH126 AND MATH300) OR MATH136}}.
-$$
-As previously stated, we denote the set of all rules for all classes as $P$.
+To solve this issues, we introduce a set of rules for each course. Each rule is
+a boolean expression represented in propositional logic in disjunctive normal
+form. For example, the rule assoicated with $MATH327$ is $$ \text{{\it MATH327:
+(MATH126 AND MATH300) OR MATH136}}.  $$ As previously stated, we denote the set
+of all rules for all classes as $P$.
 
-Furthermore, to determine if a given student is eligible to register for a course offering, we can evaluate the rule for the course, where each clause takes on the value true if a student has already taken the course, and false if he or she hasn't.
-If the rule evaluates to true, then the student is eligible to take the course; otherwise, the student is not.
+Furthermore, to determine if a given student is eligible to register for
+a course offering, we can evaluate the rule for the course, where each clause
+takes on the value true if a student has already taken the course, and false if
+he or she hasn't. If the rule evaluates to true, then the student is eligible to
+take the course; otherwise, the student is not.
 
-\subsection{Initial Model and Tracability}
-With a static schedule it would be possible to deterministically plan an entire schedule from start to finish, unfortunately we have neither the computational resources nor the luxury of a static schedule.
-From the student’s point of view, although there are some classes that are offered every quarter, senior electives are not, and sometimes it cannot be predicted more than one quarter in advance when a senior elective will be offered.
-Even given perfect information, where we know the possible class schedules for every quarter we face another problem; finding the best possible schedule reduces to enumerating possible class schedules for every quarter until the graduation requirements are satisfied.
-Given that we cannot predict in advance the conflicts between classes, we cannot even determine the branching factor of the enumeration tree nor its depth.
-Since we can only determine one quarter in advance, this suggests approaching the problem by using heuristics to pick the best possible classes each quarter.
+\subsection{Initial Model and Tracability} With a static schedule it would be
+possible to deterministically plan an entire schedule from start to finish,
+unfortunately we have neither the computational resources nor the luxury of
+a static schedule. From the student’s point of view, although there are some
+classes that are offered every quarter, senior electives are not, and sometimes
+it cannot be predicted more than one quarter in advance when a senior elective
+will be offered. Even given perfect information, where we know the possible
+class schedules for every quarter we face another problem; finding the best
+possible schedule reduces to enumerating possible class schedules for every
+quarter until the graduation requirements are satisfied. Given that we cannot
+predict in advance the conflicts between classes, we cannot even determine the
+branching factor of the enumeration tree nor its depth. Since we can only
+determine one quarter in advance, this suggests approaching the problem by using
+heuristics to pick the best possible classes each quarter.
 
-Instead of trying to solve the complete problem we chose to use the rolling horizon approach used by Xiangtong et al \cite{xiangton:informs} when scheduling classes for airlines.
-Our problem already has a well defined horizon demarcation, the quarters.
-For the rest of the paper we focus on developing methods for attempting to optimally pick classes in each quarter separately.
+Instead of trying to solve the complete problem we chose to use the rolling
+horizon approach used by Xiangtong et al \cite{xiangton:informs} when scheduling
+classes for airlines. Our problem already has a well defined horizon
+demarcation, the quarters. For the rest of the paper we focus on developing
+methods for attempting to optimally pick classes in each quarter separately.
 
-\section{Rolling Horizons Approach}
-We adapt a rolling horizons approach which uses a heuristic to pick the best possible schedule for any given quarter, given the classes offered during the quarter and the courses the student has previously taken.
-We use a greedy algorithm that functions on a per-quarter basis to accomplish this.
-The greedy algorithm greatly improves the tracability of course scheduling over an initially unknown number of quarters.
+\section{Rolling Horizons Approach} We adapt a rolling horizons approach which
+uses a heuristic to pick the best possible schedule for any given quarter, given
+the classes offered during the quarter and the courses the student has
+previously taken. We use a greedy algorithm that functions on a per-quarter
+basis to accomplish this. The greedy algorithm greatly improves the tracability
+of course scheduling over an initially unknown number of quarters.
 
-We break our algorithm down into two main phases: determining available course offering and selecting specific course offerings.
+We break our algorithm down into two main phases: determining available course
+offering and selecting specific course offerings.
 
-\subsection{Determining Available Course Offerings}
-We need to select the classes we can currently take and then build conflict graphs for classes based on the times they are offered.
-In this section, we let $L$ be the initially empty set of course offerings a given student is eligible to take in the given quarter, $q$.
-\\
+\subsection{Determining Available Course Offerings} We need to select the
+classes we can currently take and then build conflict graphs for classes based
+on the times they are offered. In this section, we let $L$ be the initially
+empty set of course offerings a given student is eligible to take in the given
+quarter, $q$. \\
 
-\subsubsection{Procedue}
-We present the procedure for popluating the set $L$ as follows:
-\begin{enumerate}
-\item For each rule in $P$ we iterate over each proposition in the rule, $p$, and check if $p \in T$.
-If $p \in T$ then we say $p$ evaluates to true; otherwise, $p$ evaluates to false.
-We recall that each rule is associated with a given course.
-If the entire rule evaluates to true, then we add the course associated with it to $L$.
-Once this is complete, $L$ will consist of every course that could be taken based solely on prerequisites that we have fulfilled.
-\item We now prune this set by removing all courses that are not offered in the current quarter.
-$L$ will necessarily now only contain classes that we can take given the current quarter and the classes we have already taken.
-\item We now construct a confict graph, $G_{conflict}$, where every course offering in $L$ is a node, and there is an edge between two nodes $i$ and $j$ if and only if the time of course offering $i$ overlaps the time of course offering $j$.
-\end{enumerate}
+\subsubsection{Procedue} We present the procedure for popluating the set $L$ as
+follows: \begin{enumerate} \item For each rule in $P$ we iterate over each
+proposition in the rule, $p$, and check if $p \in T$. If $p \in T$ then we say
+$p$ evaluates to true; otherwise, $p$ evaluates to false. We recall that each
+rule is associated with a given course. If the entire rule evaluates to true,
+then we add the course associated with it to $L$. Once this is complete, $L$
+will consist of every course that could be taken based solely on prerequisites
+that we have fulfilled.  \item We now prune this set by removing all courses
+that are not offered in the current quarter. $L$ will necessarily now only
+contain classes that we can take given the current quarter and the classes we
+have already taken.  \item We now construct a confict graph, $G_{conflict}$,
+where every course offering in $L$ is a node, and there is an edge between two
+nodes $i$ and $j$ if and only if the time of course offering $i$ overlaps the
+time of course offering $j$.  \end{enumerate}
 
-\subsubsection{Analysis}
-The selection procedure involves a lot of set queries and iteration, but in asymptotic running time the algorithm is relatively quick.
-Let $n$ be the total number of classes ever available to take, or in other words, $n = \|U\|$.
-In the worst case, $P$ has a complex rule for all $n$ courses, and checking each rule in $P$ is linear in the number of propositions.
-However, we know {\it a priori} from analysis of our dataset that the largest rule has twelve propositions, so we can bound the rule check time by a constant, $C$, and therefore step one can be done in $O(Cn)$ time.
+\subsubsection{Analysis} The selection procedure involves a lot of set queries
+and iteration, but in asymptotic running time the algorithm is relatively quick.
+Let $n$ be the total number of classes ever available to take, or in other
+words, $n = \|U\|$. In the worst case, $P$ has a complex rule for all $n$
+courses, and checking each rule in $P$ is linear in the number of propositions.
+However, we know {\it a priori} from analysis of our dataset that the largest
+rule has twelve propositions, so we can bound the rule check time by a constant,
+$C$, and therefore step one can be done in $O(Cn)$ time.
 
-Continuing to step two, the worst case is when $L$ = $Q$.
-Checking whether a course is offered in a given quarter can be done in constant time, and we have $n$ courses to check, therefore step two can also be done in $O(Cn)$ time.
+Continuing to step two, the worst case is when $L$ = $Q$. Checking whether
+a course is offered in a given quarter can be done in constant time, and we have
+$n$ courses to check, therefore step two can also be done in $O(Cn)$ time.
 
-Finally, the last step is once again bounded by a constant and $n$.
-Considering courses between 8:30am and 7:30pm, we have nine fifty-minute time periods, and five days of the week, and therefore we have 45 time slots to check in our conflict graph, again yielding $O(Cn)$ runtime.
+Finally, the last step is once again bounded by a constant and $n$. Considering
+courses between 8:30am and 7:30pm, we have nine fifty-minute time periods, and
+five days of the week, and therefore we have 45 time slots to check in our
+conflict graph, again yielding $O(Cn)$ runtime.
 
-Since we are dealing with asymptotic running time we will drop the constant terms and combining all resulsts yields an overall runtime of $O(n)$.
+Since we are dealing with asymptotic running time we will drop the constant
+terms and combining all resulsts yields an overall runtime of $O(n)$.
 
-\subsection{Selection of Quarterly Classes}
-\subsubsection{Selection Prodcedure}
-At this point in time, we know which course offerings are available for the given student to register for, and we must now pick the most optimal ones.
-There a several ways we could view the problem of determining which classes to take each quarter.
-We will initially encode the class selection as a binary integer programming problem in spirit of the methods presented by Dinkel et al \cite{dinel:scheduling}.
-and Pritsker et al \cite{prisker:informs}.
-At this point we have the set $L$ containing possible classes we that we need to take.
+\subsection{Selection of Quarterly Classes} \subsubsection{Selection Prodcedure}
+At this point in time, we know which course offerings are available for the
+given student to register for, and we must now pick the most optimal ones. There
+a several ways we could view the problem of determining which classes to take
+each quarter. We will initially encode the class selection as a binary integer
+programming problem in spirit of the methods presented by Dinkel et al
+\cite{dinel:scheduling}. and Pritsker et al \cite{prisker:informs}. At this
+point we have the set $L$ containing possible classes we that we need to take.
 
-\subsubsection{Heuristic Value Function}
-In order to determine which classes should be picked in a given schedule, we use a heuristic function which assigns each course 
-offering a value. The higher the value of this function for a given course offering is, the more likely the given course offering 
-will be placed in the schedule. Recalling that our goal is to minimize the total number of quarters taken (and implicitly, 
-maximize the total number of courses taken per quarter), our heuristic function should clearly give higher value to course 
-offerings that will help us achieve this goal. One component to the heuristic function then must be the number of dependent 
-courses a given course offering has. If a course satisfies many prerequisties, then we know that there are many courses that 
-we cannot complete without it, so we want our heuristic function to favor it.
+\subsubsection{Heuristic Value Function} In order to determine which classes
+should be picked in a given schedule, we use a heuristic function which assigns
+each course offering a value. The higher the value of this function for a given
+course offering is, the more likely the given course offering will be placed in
+the schedule. Recalling that our goal is to minimize the total number of
+quarters taken (and implicitly, maximize the total number of courses taken per
+quarter), our heuristic function should clearly give higher value to course
+offerings that will help us achieve this goal. One component to the heuristic
+function then must be the number of dependent courses a given course offering
+has. If a course satisfies many prerequisties, then we know that there are many
+courses that we cannot complete without it, so we want our heuristic function to
+favor it.
 
-Additionally, we know that not every course is offered every quarter. If a course is not offered until the following year, we 
-could potentially add an entire year to the schedule. This is a situation we hope to avoid. Our heuristic function must 
-therefore consider how often and how far in the future a given course is offered. For example, if we are deciding between 
-two course offering that are identical except one is ofference once a year, and the other is offered three times a year, we 
-want to take the course only offered once a year, as we won't have the oppurtunity to take it for another entire year.
+Additionally, we know that not every course is offered every quarter. If
+a course is not offered until the following year, we could potentially add an
+entire year to the schedule. This is a situation we hope to avoid. Our heuristic
+function must therefore consider how often and how far in the future a given
+course is offered. For example, if we are deciding between two course offering
+that are identical except one is ofference once a year, and the other is offered
+three times a year, we want to take the course only offered once a year, as we
+won't have the oppurtunity to take it for another entire year.
 
-We these considerations in mind, we begin to develop our heuristic function. In order to define our heuristic function, we 
-must first define some auxilary functions on course offering.
-\begin{itemize}
-\item Let $TimeFromNow(c, q)$ be difference in quarters between the current quarter, $q$, and the next quarter a given course offering, $c$, is offered.
-As there are four quarters, the range of this function is $\{0, 1, 2, 3\}$.
-\item Let $Quarters(c)$ be the number of quarters per year a given course, $c$, is offered. As there are four quarters, the range of this function is $\{0, 1, 2, 3, 4\}$.
-\item Let $Dependents(c)$ be the number of course that either directly or indirectly require a given course, $c$. In our final 
-draft, we will present a diagram illustrating this value more clearly.
-\end{itemize}
+We these considerations in mind, we begin to develop our heuristic function. In
+order to define our heuristic function, we must first define some auxilary
+functions on course offering.  \begin{itemize} \item Let $TimeFromNow(c, q)$ be
+difference in quarters between the current quarter, $q$, and the next quarter
+a given course offering, $c$, is offered. As there are four quarters, the range
+of this function is $\{0, 1, 2, 3\}$.  \item Let $Quarters(c)$ be the number of
+quarters per year a given course, $c$, is offered. As there are four quarters,
+the range of this function is $\{0, 1, 2, 3, 4\}$.  \item Let $Dependents(c)$ be
+the number of course that either directly or indirectly require a given course,
+$c$. In our final draft, we will present a diagram illustrating this value more
+clearly.  \end{itemize}
 
@@
 With these functions defined, we can now define our heuristic function:
-\begin{equation}
-Value(c) = 1 + \alpha * Dependents(c) + \beta * TimeFromNow(c, q) + \gamma * Quarters(c)
-\end{equation}
-Where $c$ is a given course offering, $q$ is the current quarter, and $\alpha$, $\beta$, and $\gamma$ are real numbers, giving a 
-weight to piece of the heuristic. The scalars are determined experimentally.
+\begin{equation} Value(c) = 1 + \alpha * Dependents(c) + \beta * TimeFromNow(c,
+q) + \gamma * Quarters(c) \end{equation} Where $c$ is a given course offering,
+$q$ is the current quarter, and $\alpha$, $\beta$, and $\gamma$ are real
+numbers, giving a weight to piece of the heuristic. The scalars are determined
+experimentally.
 
-\subsubsection{Formulation as Binary Integer Program}
-We can now formulate our problem as a binary integer linear programming probelm. Before stating the linear program, we must first define some variables used in it: \\
-\begin{itemize}
-\item $
-  X_i = \left\{
-     \begin{array}{lr}
-       1 \text{ if we take } X_i\\
-       0 \text{ if we don't take } X_i
-     \end{array}
-   \right. $
-\item For all $t \in O$, we let $t = 1$.
-\item Let $m = \|L\|$.
-\end{itemize}
-We now state our problem as follows:\\
-Maximize:
-$$
-\text{Value of Quarter} = \sum_{i=1}^m Value(c_i) * X_i
-$$
-Subject To:
-$$
-\sum_{i=1}^m Weight(c_i) * X_i \leq l,
-$$
-$$
-\forall t_j \in O, \sum_{i=1}^m t_j * X_i \leq 1,
-$$
-$$
-X_i \in \{0, 1\}
-$$
-The linear program can be interpreted as follows:
-\begin{itemize}
-\item The value of a quarter is the total number of courses taken that quarter.
-\item In each quarter, the total number of credits cannot exceed the specified limit.
-\item Only one course can be taken in one particular time slot
-\item A course can only be taken or not taken. In other words, decision variables are binary.
-\item Each non-zero descision variable in the optimal solution represents a class we will take.
-\end{itemize}
+\subsubsection{Formulation as Binary Integer Program} We can now formulate our
+problem as a binary integer linear programming probelm. Before stating the
+linear program, we must first define some variables used in it: \\
+\begin{itemize} \item $ X_i = \left\{ \begin{array}{lr} 1 \text{ if we take
+} X_i\\ 0 \text{ if we don't take } X_i \end{array} \right. $ \item For all $t
+\in O$, we let $t = 1$.  \item Let $m = \|L\|$.  \end{itemize} We now state our
+problem as follows:\\ Maximize: $$ \text{Value of Quarter} = \sum_{i=1}^m
+Value(c_i) * X_i $$ Subject To: $$ \sum_{i=1}^m Weight(c_i) * X_i \leq l, $$ $$
+\forall t_j \in O, \sum_{i=1}^m t_j * X_i \leq 1, $$ $$ X_i \in \{0, 1\} $$ The
+linear program can be interpreted as follows: \begin{itemize} \item The value of
+a quarter is the total number of courses taken that quarter.  \item In each
+quarter, the total number of credits cannot exceed the specified limit.  \item
+Only one course can be taken in one particular time slot \item A course can only
+be taken or not taken. In other words, decision variables are binary.  \item
+Each non-zero descision variable in the optimal solution represents a class we
+will take.  \end{itemize}
 
-It is worth nothing that all of the t variables are essentially dummy variables that ensure the correct number of equations of type 2.
-Each of the equations based on eqn.
-2 ensure that we only take one class per time period per day.
-Also this linear program is ‘nice’ in that it always has feasible origin (don’t take any classes), and that the simplest formulation is already in standard form.
+It is worth nothing that all of the t variables are essentially dummy variables
+that ensure the correct number of equations of type 2. Each of the equations
+based on eqn. 2 ensure that we only take one class per time period per day. Also
+this linear program is ‘nice’ in that it always has feasible origin (don’t take
+any classes), and that the simplest formulation is already in standard form.
 
-As noted in the reference literature this is actually a (strongly) NP-complete problem as described \cite{pferschy:kcg}.
-For the rest of this section we consider an alternative formulation of class selection as a knapsack problem with conflict graphs and suggest the use a technique developed by Yamada et.
-al.
+As noted in the reference literature this is actually a (strongly) NP-complete
+problem as described \cite{pferschy:kcg}. For the rest of this section we
+consider an alternative formulation of class selection as a knapsack problem
+with conflict graphs and suggest the use a technique developed by Yamada et. al.
 \cite{yamada:heuristic}.
 
-We begin by simplifying $G_{conflict}$ to be a fully connected graph by adding dummy nodes $v'$ and insert edges that connect $v'$ to every other node in the graph.
-To ensure that $v'$ does not alter the final optimal solution, we assign it the following properties:
-\begin{itemize}
-\item $Weight(v') = l+1$
-\item $Value(v') = 0$
-\end{itemize}
-As the weight of our dummy node is greater than our credit limit, and it's value is zero, we know that it will never appear in the optimal solution.
-With the above reasoning we show for our case of class schedule that Pferschy and Schauer's \cite{pferschy:kcg} claim that adding a dummy node without loss of generality holds.
-We can now conveniently write the class picking problem as the knapsack with conflict graphs problem (KCG): \\
-Maximize:
-$$
-\text{Value of Quarter} = \sum_{i=1}^m Value(c_i) * X_i
-$$
-Subject To:
-$$
-\sum_{i=1}^m Weight(c_i) * X_i \leq l,
-$$
-$$
-X_i + X_j \leq 1 \text{ for all connected vertices } i, j \in G_{conflict},
-$$
-$$
-X_i \in \{0, 1\}
-$$
-It is worth noting that for relatively small versions of this problem, in the low hundreds of variables it is possible to solve this problem exactly using modern LP solving software \cite{yamada:heuristic}.
-In our case, just solving for the optimal classes for a single major fits well into the scope of a few hundred (or even just tens of) variables.
-For multiple majors or to consider class selection for distribution and elective requirements, however the number of classes balloons significantly and the KCG problem becomes much harder.
-For this version of the problem we suggest using the techniques used by Yamada et al \cite{yamada:heuristic}, which we briefly outline below.
+We begin by simplifying $G_{conflict}$ to be a fully connected graph by adding
+dummy nodes $v'$ and insert edges that connect $v'$ to every other node in the
+graph. To ensure that $v'$ does not alter the final optimal solution, we assign
+it the following properties: \begin{itemize} \item $Weight(v') = l+1$ \item
+$Value(v') = 0$ \end{itemize} As the weight of our dummy node is greater than
+our credit limit, and it's value is zero, we know that it will never appear in
+the optimal solution. With the above reasoning we show for our case of class
+schedule that Pferschy and Schauer's \cite{pferschy:kcg} claim that adding
+a dummy node without loss of generality holds. We can now conveniently write the
+class picking problem as the knapsack with conflict graphs problem (KCG): \\
+Maximize: $$ \text{Value of Quarter} = \sum_{i=1}^m Value(c_i) * X_i $$ Subject
+To: $$ \sum_{i=1}^m Weight(c_i) * X_i \leq l, $$ $$ X_i + X_j \leq 1 \text{ for
+all connected vertices } i, j \in G_{conflict}, $$ $$ X_i \in \{0, 1\} $$ It is
+worth noting that for relatively small versions of this problem, in the low
+hundreds of variables it is possible to solve this problem exactly using modern
+LP solving software \cite{yamada:heuristic}. In our case, just solving for the
+optimal classes for a single major fits well into the scope of a few hundred (or
+even just tens of) variables. For multiple majors or to consider class selection
+for distribution and elective requirements, however the number of classes
+balloons significantly and the KCG problem becomes much harder. For this version
+of the problem we suggest using the techniques used by Yamada et al
+\cite{yamada:heuristic}, which we briefly outline below.
 
-\subsubsection{Relaxation and Implicit Enumeration}
-\begin{enumerate}
-\item Develop upper bound on value by solving a continuous version knapsack problem generated by a Lagrangean relaxation of our problem.
-Let $X_i$ be a continuous variable with $0 \leq X_i \leq 1$, and shift the time constraints into the objective function by using a penalty of $\lambda$ when the constraint is violated.
-\item Use the upper bound developed above with the pruning rules suggested by Yamada et al \cite{yamada:heuristic} to do implicit enumeration of the possible choices.
-\end{enumerate}
+\subsubsection{Relaxation and Implicit Enumeration} \begin{enumerate} \item
+Develop upper bound on value by solving a continuous version knapsack problem
+generated by a Lagrangean relaxation of our problem. Let $X_i$ be a continuous
+variable with $0 \leq X_i \leq 1$, and shift the time constraints into the
+objective function by using a penalty of $\lambda$ when the constraint is
+violated.  \item Use the upper bound developed above with the pruning rules
+suggested by Yamada et al \cite{yamada:heuristic} to do implicit enumeration of
+the possible choices.  \end{enumerate}
 
-In Yamada et al’s \cite{yamada:heuristic} numerical runtime analysis they found this algorithm ran very favorably against solving the same problem using popular linear programming solvers.
-There is one problem that we should briefly address, because the number of periods of available classes per week is relatively small the conflict rate for a large number of class is relatively higher than Yamada et al’s 2\% maximal conflict rate for their experiments.
-Our data has a significant mitigating factor though, knapsack problems in general are exponential in terms of the number of choices and the total weight capacity.
-Realistically the total weight capacity is going to be 20 credits or less which is between 1\% and 2\% of the figures used by Yamada \cite{yamada:heuristic} and also our total number of choices is in many cases (particular just the classes in a particular major) less than 200.
-Giving the size and type of data in our set the higher conflict rate is less of a concern and Yamada et al’s algorithm should still be efficient.
+In Yamada et al’s \cite{yamada:heuristic} numerical runtime analysis they found
+this algorithm ran very favorably against solving the same problem using popular
+linear programming solvers. There is one problem that we should briefly address,
+because the number of periods of available classes per week is relatively small
+the conflict rate for a large number of class is relatively higher than Yamada
+et al’s 2\% maximal conflict rate for their experiments. Our data has
+a significant mitigating factor though, knapsack problems in general are
+exponential in terms of the number of choices and the total weight capacity.
+Realistically the total weight capacity is going to be 20 credits or less which
+is between 1\% and 2\% of the figures used by Yamada \cite{yamada:heuristic} and
+also our total number of choices is in many cases (particular just the classes
+in a particular major) less than 200. Giving the size and type of data in our
+set the higher conflict rate is less of a concern and Yamada et al’s algorithm
+should still be efficient.
 
-\section{Computational Results}
-In this section we will outline results obtained by running computer simulations of schedule picking.
-We will computationally examine what parameters in our heuristic function produce the most desirable schedules.
+\section{Computational Results} In this section we will outline results obtained
+by running computer simulations of schedule picking. We will computationally
+examine what parameters in our heuristic function produce the most desirable
+schedules.
 
-\bibliography{references}{}
-\bibliographystyle{plain}
-\end{document}
+\bibliography{references}{} \bibliographystyle{plain} \end{document}

### 2010-12-04 paper.tex
Now spell checked

@@
 \section{Introduction} \subsection{Background} Registration can be a stressful
-time for every student at the Univeristy of Washington. Without deliberate
+time for every student at the University of Washington. Without deliberate
 planning, students must decide which classes to take on a quarterly basis. The
@@
 that minimizes the number of quarters until graduation, eliminates time
-conflicts, and eliminates prerequiste conflicts.
+conflicts, and eliminates prerequisite conflicts.
 
@@
 the student's current state, which schedules are the most ideal in terms of
-saving time and resources. This will also help studentssave money, as they will
+saving time and resources. This will also help students save money, as they will
 be in school for fewer quarters. The university also benefits from the reduced
@@
 possibly ever take.  \item Let $P$ be a set of rules specifying course
-prerequistes.  \item Let $l$ be the maximum number of credits a student wishes
+prerequisites.  \item Let $l$ be the maximum number of credits a student wishes
 to take during any given quarter.  \item Let $O$ be the set of all 50-minute
@@
 \subsection{Representing Prerequisites} Initially, we hoped to represent course
-prerequistes using a directed acyclic graph, in which each node was a course,
+prerequisites using a directed acyclic graph, in which each node was a course,
 and edges between nodes $i$ and $j$ if and only if $i$ was a prerequisite for
@@
 a boolean expression represented in propositional logic in disjunctive normal
-form. For example, the rule assoicated with $MATH327$ is $$ \text{{\it MATH327:
+form. For example, the rule associated with $MATH327$ is $$ \text{{\it MATH327:
 (MATH126 AND MATH300) OR MATH136}}.  $$ As previously stated, we denote the set
@@
 
-\subsection{Initial Model and Tracability} With a static schedule it would be
+\subsection{Initial Model and Tractability} With a static schedule it would be
 possible to deterministically plan an entire schedule from start to finish,
@@
 previously taken. We use a greedy algorithm that functions on a per-quarter
-basis to accomplish this. The greedy algorithm greatly improves the tracability
+basis to accomplish this. The greedy algorithm greatly improves the tractability
 of course scheduling over an initially unknown number of quarters.
@@
 
-\subsubsection{Procedue} We present the procedure for popluating the set $L$ as
+\subsubsection{Procedure} We present the procedure for populating the set $L$ as
 follows: \begin{enumerate} \item For each rule in $P$ we iterate over each
@@
 contain classes that we can take given the current quarter and the classes we
-have already taken.  \item We now construct a confict graph, $G_{conflict}$,
+have already taken.  \item We now construct a conflict graph, $G_{conflict}$,
 where every course offering in $L$ is a node, and there is an edge between two
@@
 courses, and checking each rule in $P$ is linear in the number of propositions.
-However, we know {\it a priori} from analysis of our dataset that the largest
+However, we know {\it a priori} from analysis of our data set that the largest
 rule has twelve propositions, so we can bound the rule check time by a constant,
@@
 Since we are dealing with asymptotic running time we will drop the constant
-terms and combining all resulsts yields an overall runtime of $O(n)$.
+terms and combining all results yields an overall run time of $O(n)$.
 
-\subsection{Selection of Quarterly Classes} \subsubsection{Selection Prodcedure}
+\subsection{Selection of Quarterly Classes} \subsubsection{Selection Procedure}
 At this point in time, we know which course offerings are available for the
@@
 function then must be the number of dependent courses a given course offering
-has. If a course satisfies many prerequisties, then we know that there are many
+has. If a course satisfies many prerequisites, then we know that there are many
 courses that we cannot complete without it, so we want our heuristic function to
@@
 course is offered. For example, if we are deciding between two course offering
-that are identical except one is ofference once a year, and the other is offered
+that are identical except one is offered once a year, and the other is offered
 three times a year, we want to take the course only offered once a year, as we
-won't have the oppurtunity to take it for another entire year.
+won't have the opportunity to take it for another entire year.
 
 We these considerations in mind, we begin to develop our heuristic function. In
-order to define our heuristic function, we must first define some auxilary
+order to define our heuristic function, we must first define some auxiliary
 functions on course offering.  \begin{itemize} \item Let $TimeFromNow(c, q)$ be
@@
 \subsubsection{Formulation as Binary Integer Program} We can now formulate our
-problem as a binary integer linear programming probelm. Before stating the
+problem as a binary integer linear programming problem. Before stating the
 linear program, we must first define some variables used in it: \\
@@
 be taken or not taken. In other words, decision variables are binary.  \item
-Each non-zero descision variable in the optimal solution represents a class we
+Each non-zero decision variable in the optimal solution represents a class we
 will take.  \end{itemize}
@@
 Develop upper bound on value by solving a continuous version knapsack problem
-generated by a Lagrangean relaxation of our problem. Let $X_i$ be a continuous
+generated by a Lagrangian relaxation of our problem. Let $X_i$ be a continuous
 variable with $0 \leq X_i \leq 1$, and shift the time constraints into the

### 2010-12-04 paper.tex
Fixed some of the references and scaled the graph figure by 0.5

@@
 planning, students must decide which classes to take on a quarterly basis. The
-decision process involves considering many factors, such as quarterly class
+decision process involves considering factors such as quarterly class
 offerings, the amount of credits, the prerequisites of the classes, time
@@
 programming problem in spirit of the methods presented by Dinkel et al
-\cite{dinel:scheduling}. and Pritsker et al \cite{prisker:informs}. At this
+\cite{dinkel:scheduling}. and Pritsker et al \cite{prisker:informs}. At this
 point we have the set $L$ containing possible classes we that we need to take.
@@
 
-\includegraphics{more_prereq_tree}
+\includegraphics[scale=0.5]{more_prereq_tree}
 
@@
 
-\bibliography{references}{} \bibliographystyle{plain} \end{document}
+\bibliography{paper_references}{} \bibliographystyle{plain} \end{document}

### 2010-12-04 paper/paper.tex
moved all paper files to a seperate subdirectory, added new figure data file and added figure to paper in 2.2

--- 2010-12-04 paper.tex
@@
 \usepackage{paralist} % very flexible & customisable lists (eg. enumerate/itemize, etc.)
+\usepackage{epsfig}
 \usepackage{verbatim} % adds environment for commenting out blocks of text & for better verbatim
@@
 
+\begin{figure} [ht]
+    \begin{center}
+        \includegraphics[scale=0.4]{prereq_logic_example}
+    \end{center}
+    \caption{Evaluation of a rule}
+    \label{logic_ex}
+\end{figure}
+
 To solve this issues, we introduce a set of rules for each course. Each rule is
@@
 he or she hasn't. If the rule evaluates to true, then the student is eligible to
-take the course; otherwise, the student is not.
+take the course; otherwise, the student is note. For example, consider
+figure~\ref{logic_ex}, here a student has previously taken courses $MATH126$ and
+$MATH300$. Since $MATH126$ and $MATH300$ have been taken, then those literals in
+the rule for $MATH327$ evaluate to true and therefore the entire rule is true.
+Since the entire rule is true, the student can take $MATH327$ when next
+available.
 
@@
 
+\begin{figure} [ht]
+    \begin{center}
+        \includegraphics[scale=0.35]{more_prereq_tree}
+    \end{center}
+    \caption{A Prerequisite DAG}
+    \label{prereq}
+\end{figure}
+
 We these considerations in mind, we begin to develop our heuristic function. In
@@
 the number of course that either directly or indirectly require a given course,
-$c$. In our final draft, we will present a diagram illustrating this value more
-clearly.  \end{itemize}
-
-\includegraphics[scale=0.5]{more_prereq_tree}
+$c$. In Figure~\ref{prereq}, $Dependents(MATH125)$ = 6 while
+$Dependents(MATH307)$ = 1 \end{itemize}
 

### 2010-12-05 paper/paper.tex
fiddled with paper type and margin settings slightly
proof read, chopped out a couple sentences, fixed some awkward ones, added in a couple places

@@
 \usepackage{geometry} % to change the page dimensions
-\geometry{a4paper} % or letterpaper (US) or a5paper or....
-% \geometry{margins=2in} % for example, change the margins to 2 inches all round
-% \geometry{landscape} % set up the page for landscape
-%   read geometry.pdf for detailed page layout information
+\geometry{top=3cm,bottom=3cm,left=3cm,right=3cm,letterpaper} % or letterpaper (US) or a5paper or....
 
@@
 however, it only takes into account time conflicts, and does not give advice on
-how to pick the most ideal schedule. The situation described above makes it
-difficult for students advising faculties. A student's graduation time may be
-extended because of his or her inability to take a class that has limited
-offerings during that particular year, for reasons such as not satisfying the
-prerequisites of the class. The advising office’s workload can also be increased
-due to the changing plans and majors of students and their need for repeated
-advising.
+how to pick the most ideal schedule.
+
+A student's graduation time may be extended because of his or her inability to
+take a class that has limited offerings during that particular year.
+Additionally, the advising office’s workload can also be increased due to the
+changing plans and majors of students and their need for repeated advising.
 
@@
 school’s registration system will benefit the students by enabling them to map
-out their quarterly class schedules until graduation, and to preview their ideal
-schedule, the length of time until graduation, and the credit hours per quarter.
-This will lead to better decision making for what majors to pursue, and given
-the student's current state, which schedules are the most ideal in terms of
-saving time and resources. This will also help students save money, as they will
-be in school for fewer quarters. The university also benefits from the reduced
-labor cost and increased efficiency of the advising offices.
+out their quarterly class schedules until graduation. Students will be able to
+preview their ideal schedule, the length of time until graduation, and the
+credit hours per quarter.  This will lead to better decision making for what
+majors to pursue, and given the student's current state, which schedules are the
+most ideal in terms of saving time and resources. This will also help students
+save money, as they will be in school for fewer quarters. The university also
+benefits from the reduced labor cost and increased efficiency of the advising
+offices.
 
@@
 time periods. Since there are far more classes than time periods during the
-regular five-day week there are inevitably conflicts between courses that
-a student wants to take and courses that a student needs to take. Or worse,
-there are conflicts between two or more courses that a students needs to
-graduate. To graduate with a B.S. in Mathematics there are certain classes that
-are required and then there are elective requirements too. Additionally
-a student may wish to take some classes that are not required.  Adding to the
-problem is many classes cannot just be taken whenever they are available;
+regular five-day week there are inevitably time conflicts between courses that
+a student could take. To graduate with a B.S. in Mathematics there are certain
+classes that are required and then there are elective requirements too. Adding
+to the problem is many classes cannot just be taken whenever they are available;
 instead, certain prerequisites must be satisfied first. In the case of math
@@
 
-We must now also define some variables used throughout our paper.
-\begin{itemize} \item Let $S$ be the set of courses required to be taken in
-order to graduate. $S$ can be thought of as our goal state, and can also contain
-elective courses a student simply wishes to take.  \item Let $T$ be the set of
-courses a given student has already completed. If no courses have been taken,
-this set may be empty.  \item Let $U$ be the set of all courses a student could
-possibly ever take.  \item Let $P$ be a set of rules specifying course
-prerequisites.  \item Let $l$ be the maximum number of credits a student wishes
-to take during any given quarter.  \item Let $O$ be the set of all 50-minute
-time blocks between 8:30am and 7:30pm from Monday through Friday.  \end{itemize}
+Additionally: \begin{itemize} \item Let $S$ be the set of courses required to be
+taken in order to graduate. $S$ can be thought of as our goal state, and can
+also contain elective courses a student simply wishes to take.  \item Let $T$ be
+the set of courses a given student has already completed. If no courses have
+been taken, this set may be empty.  \item Let $U$ be the set of all courses
+a student could possibly ever take.  \item Let $P$ be a set of rules specifying
+course prerequisites.  \item Let $l$ be the maximum number of credits a student
+wishes to take during any given quarter.  \item Let $O$ be the set of all
+50-minute time blocks between 8:30am and 7:30pm from Monday through Friday.
+\end{itemize}
 
 \subsection{Representing Prerequisites} Initially, we hoped to represent course
-prerequisites using a directed acyclic graph, in which each node was a course,
-and edges between nodes $i$ and $j$ if and only if $i$ was a prerequisite for
-$j$.  This method was used by Pritsker et. al. and Xiangtong et. al. However, it
-does not work in our model, and courses can have a complicated set of multiple
-prerequisites. For example, {\it MATH327} requires both $MATH126$ and $MATH300$,
-or just $MATH136$. Let this graph be called $G_{prereq}$. Due to these
-complications, a $G_{prereq}$ alone will not suffice.
+prerequisites using a directed acyclic graph. Each node was a course, and edges
+existed between nodes $i$ and $j$ if and only if $i$ was a prerequisite for $j$.
+However, it does not work in our model, and courses can have a complicated set
+of multiple prerequisites.  For example, {\it MATH327} requires both $MATH126$
+and $MATH300$, or just $MATH136$.
 
@@
 \caption{Evaluation of a rule} 
-    \label{logic_ex}
-\end{figure}
+\label{logic_ex} \end{figure}
 
@@
 unfortunately we have neither the computational resources nor the luxury of
-a static schedule. From the student’s point of view, although there are some
-classes that are offered every quarter, senior electives are not, and sometimes
-it cannot be predicted more than one quarter in advance when a senior elective
-will be offered. Even given perfect information, where we know the possible
-class schedules for every quarter we face another problem; finding the best
-possible schedule reduces to enumerating possible class schedules for every
+a static schedule to do plan. From the student’s point of view, although there
+are some classes that are offered every quarter, senior electives are not, and
+sometimes it cannot be predicted more than one quarter in advance when a senior
+elective will be offered. Even given perfect information, where we know the
+possible class schedules for every quarter we face another problem; finding the
+best possible schedule reduces to enumerating possible class schedules for every
 quarter until the graduation requirements are satisfied. Given that we cannot
@@
 demarcation, the quarters. For the rest of the paper we focus on developing
-methods for attempting to optimally pick classes in each quarter separately.
+methods to optimally pick classes in each quarter separately.
 
@@
 empty set of course offerings a given student is eligible to take in the given
-quarter, $q$. \\
+quarter, $q$.
 
@@
 Since we are dealing with asymptotic running time we will drop the constant
-terms and combining all results yields an overall run time of $O(n)$.
+terms and combining all results yields an overall runtime of $O(n)$.
 
@@
 At this point in time, we know which course offerings are available for the
-given student to register for, and we must now pick the most optimal ones. There
-a several ways we could view the problem of determining which classes to take
-each quarter. We will initially encode the class selection as a binary integer
-programming problem in spirit of the methods presented by Dinkel et al
-\cite{dinkel:scheduling}. and Pritsker et al \cite{prisker:informs}. At this
-point we have the set $L$ containing possible classes we that we need to take.
+given student to register for, and we must now pick the most optimal ones.  We
+present two ways of approaching quarterly class choice. We will initially encode
+the class selection as a binary integer programming problem in spirit of the
+methods presented by Dinkel et al \cite{dinkel:scheduling}. and Pritsker et al
+\cite{prisker:informs}. A second approach developed by Yamada et al
+\cite{yamada:heuristic} and Pferschy et al \cite{pferschy:kcg} is presented for
+dealing with very large class choices where binary LP would be difficult solve.
 
@@
 quarters taken (and implicitly, maximize the total number of courses taken per
-quarter), our heuristic function should clearly give higher value to course
-offerings that will help us achieve this goal. One component to the heuristic
-function then must be the number of dependent courses a given course offering
-has. If a course satisfies many prerequisites, then we know that there are many
-courses that we cannot complete without it, so we want our heuristic function to
-favor it.
+quarter). Our heuristic function should give higher value to course offerings
+that will help us achieve this goal. One component to the heuristic function
+then must be the number of dependent courses a given course offering has. If
+a course satisfies many prerequisites, then we know that there are many courses
+that we cannot complete without it, so we want our heuristic function to favor
+it.
 
@@
 \subsubsection{Formulation as Binary Integer Program} We can now formulate our
-problem as a binary integer linear programming problem. Before stating the
-linear program, we must first define some variables used in it: \\
-\begin{itemize} \item $ X_i = \left\{ \begin{array}{lr} 1 \text{ if we take
-} X_i\\ 0 \text{ if we don't take } X_i \end{array} \right. $ \item For all $t
-\in O$, we let $t = 1$.  \item Let $m = \|L\|$.  \end{itemize} We now state our
-problem as follows:\\ Maximize: $$ \text{Value of Quarter} = \sum_{i=1}^m
-Value(c_i) * X_i $$ Subject To: $$ \sum_{i=1}^m Weight(c_i) * X_i \leq l, $$ $$
-\forall t_j \in O, \sum_{i=1}^m t_j * X_i \leq 1, $$ $$ X_i \in \{0, 1\} $$ The
-linear program can be interpreted as follows: \begin{itemize} \item The value of
-a quarter is the total number of courses taken that quarter.  \item In each
-quarter, the total number of credits cannot exceed the specified limit.  \item
-Only one course can be taken in one particular time slot \item A course can only
-be taken or not taken. In other words, decision variables are binary.  \item
-Each non-zero decision variable in the optimal solution represents a class we
-will take.  \end{itemize}
+problem as a binary linear programming problem. Before stating the linear
+program, we must first define some variables used in it: \\ \begin{itemize}
+\item $ X_i = \left\{ \begin{array}{lr} 1 \text{ if we take } X_i\\ 0 \text{ if
+we don't take } X_i \end{array} \right. $ \item For all $t \in O$, we let $t
+= 1$.  \item Let $m = \|L\|$.  \end{itemize} We now state our problem as
+follows:\\ Maximize: $$ \text{Value of Quarter} = \sum_{i=1}^m Value(c_i) * X_i
+$$ Subject To: $$ \sum_{i=1}^m Weight(c_i) * X_i \leq l, $$ $$ \forall t_j \in
+O, \sum_{i=1}^m t_j * X_i \leq 1, $$ $$ X_i \in \{0, 1\} $$ The linear program
+can be interpreted as follows: \begin{itemize} \item The value of a quarter is
+the total number of courses taken that quarter.  \item In each quarter, the
+total number of credits cannot exceed the specified limit.  \item Only one
+course can be taken in one particular time slot \item A course can only be taken
+or not taken. In other words, decision variables are binary.  \item Each
+non-zero decision variable in the optimal solution represents a class we will
+take.  \end{itemize}
 

### 2010-12-05 paper/paper.tex
fixed some weasel words

@@
 graduation that given the above properties minimizes the total number of
-quarters needed. Slightly more formally we would like to: \\ Minimize the total
+quarters needed. Explicitly: \\ Minimize the total
 number of quarters needed to graduate subject to: \begin{enumerate} \item Only

### 2010-12-05 paper/paper.tex
fixed bib errors, number equations

@@
 a boolean expression represented in propositional logic in disjunctive normal
-form. For example, the rule associated with $MATH327$ is $$ \text{{\it MATH327:
-(MATH126 AND MATH300) OR MATH136}}.  $$ As previously stated, we denote the set
-of all rules for all classes as $P$.
+form. For example, the rule associated with $MATH327$ in fig~\ref{logic_ex} is
+$$ \text{{\it MATH327: (MATH126 AND MATH300) OR MATH136}}.  $$ As previously
+stated, we denote the set of all rules for all classes as $P$.
 
@@
 With these functions defined, we can now define our heuristic function:
-\begin{equation} Value(c) = 1 + \alpha * Dependents(c) + \beta * TimeFromNow(c,
-q) + \gamma * Quarters(c) \end{equation} Where $c$ is a given course offering,
+
+\begin{equation} 
+    Value(c) = 1 + \alpha * Dependents(c) + \beta * TimeFromNow(c,q) 
+    - \gamma * Quarters(c)
+    \label{value_func}
+\end{equation}
+
+Where $c$ is a given course offering,
 $q$ is the current quarter, and $\alpha$, $\beta$, and $\gamma$ are real
@@
 = 1$.  \item Let $m = \|L\|$.  \end{itemize} We now state our problem as
-follows:\\ Maximize: $$ \text{Value of Quarter} = \sum_{i=1}^m Value(c_i) * X_i
-$$ Subject To: $$ \sum_{i=1}^m Weight(c_i) * X_i \leq l, $$ $$ \forall t_j \in
-O, \sum_{i=1}^m t_j * X_i \leq 1, $$ $$ X_i \in \{0, 1\} $$ The linear program
+follows:
+
+Maximize:
+
+\begin{equation}
+    \text{Value of Quarter} = \sum_{i=1}^m Value(c_i) * X_i
+    \label{qtr_val}
+\end{equation}
+
+Subject To:
+
+\begin{equation}
+    \sum_{i=1}^m Weight(c_i) * X_i \leq l, \forall t_j \in O,
+    \label{weight_lim}
+\end{equation}
+
+\begin{equation}
+    \sum_{i=1}^m t_j * X_i \leq 1, X_i \in \{0, 1\}
+    \label{time_con}
+\end{equation}
+
+The linear program
 can be interpreted as follows: \begin{itemize} \item The value of a quarter is
@@
 
-It is worth nothing that all of the t variables are essentially dummy variables
-that ensure the correct number of equations of type 2. Each of the equations
-based on eqn. 2 ensure that we only take one class per time period per day. Also
-this linear program is ‘nice’ in that it always has feasible origin (don’t take
-any classes), and that the simplest formulation is already in standard form.
+Each of the equations based on eqn. 2 ensure that we only take one class per 
+time period per day. Also this linear program is ‘nice’ in that it always has 
+feasible origin (don’t take any classes), and that the simplest formulation 
+is already in standard form.
 
@@
 a dummy node without loss of generality holds. We can now conveniently write the
-class picking problem as the knapsack with conflict graphs problem (KCG): \\
-Maximize: $$ \text{Value of Quarter} = \sum_{i=1}^m Value(c_i) * X_i $$ Subject
-To: $$ \sum_{i=1}^m Weight(c_i) * X_i \leq l, $$ $$ X_i + X_j \leq 1 \text{ for
-all connected vertices } i, j \in G_{conflict}, $$ $$ X_i \in \{0, 1\} $$ It is
-worth noting that for relatively small versions of this problem, in the low
-hundreds of variables it is possible to solve this problem exactly using modern
-LP solving software \cite{yamada:heuristic}. In our case, just solving for the
-optimal classes for a single major fits well into the scope of a few hundred (or
-even just tens of) variables. For multiple majors or to consider class selection
+class picking problem as the knapsack with conflict graphs problem (KCG):
+
+Maximize:
+
+\begin{equation}
+    \text{Value of Quarter} = \sum_{i=1}^m Value(c_i) * X_i
+\end{equation}
+
+Subject To:
+
+\begin{equation}
+\sum_{i=1}^m Weight(c_i) * X_i \leq l,
+\end{equation}
+
+\begin{equation}
+X_i + X_j \leq 1 \text{ for all connected vertices } i, j \in G_{conflict},
+\end{equation}
+
+\begin{equation}
+X_i \in \{0, 1\}
+\end{equation}
+
+For versions of this problem in the low hundreds of variables it is possible 
+to solve this problem exactly using modern LP solving software 
+\cite{yamada:heuristic}. In our case, just solving for the optimal classes for 
+a single major fits well into the scope of a few hundred (or even just tens of) 
+variables. For multiple majors or to consider class selection
 for distribution and elective requirements, however the number of classes

### 2010-12-06 paper/paper.tex
Fixed an equation reference

@@
 
-\subsubsection{Formulation as Binary Integer Program} We can now formulate our
-problem as a binary linear programming problem. Before stating the linear
+\subsubsection{Formulation as Binary Integer Program} We can now formulate
+a binary linear programming problem. Before stating the linear
 program, we must first define some variables used in it: \\ \begin{itemize}
@@
 
-Each of the equations based on eqn. 2 ensure that we only take one class per 
+Each of the equations based on eqn.~\ref{time_con} ensure that we only take one class per 
 time period per day. Also this linear program is ‘nice’ in that it always has 

### 2010-12-10 paper/paper.tex
made equation 4 actually make sense

@@
 a binary linear programming problem. Before stating the linear
-program, we must first define some variables used in it: \\ \begin{itemize}
+program, we must first define some variables used in it:
+
+\begin{itemize}
     \item $ X_i = \left\{ \begin{array}{lr} 1 \text{ if we take } X_i\\ 0 \text{ if
-we don't take } X_i \end{array} \right. $ \item For all $t \in O$, we let $t
-= 1$.  \item Let $m = \|L\|$.  \end{itemize} We now state our problem as
-follows:
-
+we don't take } X_i \end{array} \right. $ 
+    \item For all $t \in O$, we let $t$ represent the conflict graph of classes
+    at time period $t$.
+    \item Let $m = \|L\|$.  
+\end{itemize} 
+We now state our problem as follows: 
 Maximize:
@@
 \end{equation}
-
 Subject To:
-
 \begin{equation}
-    \sum_{i=1}^m Weight(c_i) * X_i \leq l, \forall t_j \in O,
+    \sum_{i=1}^m Weight(c_i) * X_i \leq l, 
     \label{weight_lim}
@@
 \begin{equation}
-    \sum_{i=1}^m t_j * X_i \leq 1, X_i \in \{0, 1\}
+    \forall t \in O,X_i+X_j \leq 1, \forall pairs (X_i,X_j) \in t
     \label{time_con}

### 2010-12-10 quarter_page_extracter.py
Minor change to the regex for matching class names, made it a stricter match to fix some false positives

@@ -32,7 +32,7 @@ def main(args):
     lines = open(args[1], 'r').readlines()
     classes = {}
     # prepare regular expressions
-    math_class = re.compile(r"^\*?MATH\s+(\d\d\d).*\*")
+    math_class = re.compile(r"^\*?MATH\s+(\d\d\d).*<.*>\*")
     main_section = re.compile(r"(\d{5})\s<.*>\s(\w)\s+(\d)\s+(\w{1,6})\s+(\d{3,4})-(\d{3,4})")
     qz_section = re.compile(r"(\d{5})\s<.*>\s(\w\w)\s+QZ\s+(\w{1,6})\s+(\d{3,4})-(\d{3,4})")
     curr_class = 0
@@ -43,6 +43,7 @@ def main(args):
         qz_section_match = qz_section.search(line)
         # priority one is switching the current class
         if math_class_match is not None:
+            print("Matched "+math_class_match.group(1))
             curr_class = int(math_class_match.group(1))
         elif main_section_match is not None:
             curr_sec = offering(main_section_match.group(1),

### 2010-12-11 paper/paper.tex
Added constraint to stop up from taking the same course twice in once quarter
fixed up the only take one course constraint to be in both LPs
initial rewrite of opening sections and diagram add

@@
 \usepackage{geometry} % to change the page dimensions
-\geometry{top=3cm,bottom=3cm,left=3cm,right=3cm,letterpaper} % or letterpaper (US) or a5paper or....
+\geometry{top=2cm,bottom=2cm,left=2cm,right=2cm,letterpaper} % or letterpaper (US) or a5paper or....
 
@@
 
-% Adds a table of contents. Comment out if we don't want this.
-\tableofcontents
+\abstract{The Bachelor's Degree is designed to take four years to complete,
+however, many students fail to complete it this amount of time. One 
+factor contributing to this is poor planning when registering for classes.
+We present an algorithm to assist students in
+constructing a schedule that works towards completing a degree in the shortest
+amount of time possible.}
 \pagebreak
 
-\abstract{The Bachelor's Degree is designed to take four years to complete;
-however, many students fail to complete it this amount of time. One of the
-factors contributing to this is a lack of proper planning during the
-registration phase. In this paper we present an algorithm to assist students in
-constructing a schedule that works towards completing a degree in the shortest
-amount of time possible.}
-
 \section{Introduction} \subsection{Background} Registration can be a stressful
-time for every student at the University of Washington. Without deliberate
-planning, students must decide which classes to take on a quarterly basis. The
-decision process involves considering factors such as quarterly class
-offerings, the amount of credits, the prerequisites of the classes, time
-conflicts, and how the choices in current quarter affect future quarters. The
-University of Washington registration system provides a schedule finder;
-however, it only takes into account time conflicts, and does not give advice on
-how to pick the most ideal schedule.
+time for every student at the University of Washington. Without good
+planning, students can make poor registration choices resulting in more quarter
+being needed to graduate than otherwise necessary. Picking the correct class
+involves considering quarterly class offerings, class credits, the prerequisites
+of a class, time conflicts, and how to fulfill future prerequisites. The
+University of Washington registration system already provides a tool to aid
+students, schedule finder. Schedule finder however, only takes into account time
+conflicts in one quarter, and does not consider what classes are best for
+minimizing graduation time.
 
@@
 take a class that has limited offerings during that particular year.
-Additionally, the advising office’s workload can also be increased due to the
-changing plans and majors of students and their need for repeated advising.
 
-\subsection{Goal} Our goal is to alleviate the problematic situation by creating
-a mathematical algorithm that accomplishes the following goal: given a student's
-current state and preferred credit hours per quarter, produce a class schedule
-that minimizes the number of quarters until graduation, eliminates time
-conflicts, and eliminates prerequisite conflicts.
+\subsection{Goal} Our goal is to provide an automated system that gives quarter
+by quarter schedules until a student graduates. As input our system will take in
+the current set of classes a student has taken and the preferred credit hours
+per quarter. Additional input will be the current and past university time
+schedules. As output our system will provide a class schedule that minimizes the
+number of quarters until graduation.
 
-\subsection{Importance} The implementation of the algorithm as part of the
-school’s registration system will benefit the students by enabling them to map
-out their quarterly class schedules until graduation. Students will be able to
-preview their ideal schedule, the length of time until graduation, and the
-credit hours per quarter.  This will lead to better decision making for what
-majors to pursue, and given the student's current state, which schedules are the
-most ideal in terms of saving time and resources. This will also help students
-save money, as they will be in school for fewer quarters. The university also
-benefits from the reduced labor cost and increased efficiency of the advising
-offices.
+\subsection{Importance} The implementation of our system as part of the school’s
+registration system will benefit the students by enabling them to map out their
+quarterly class schedules until graduation. Students will be able to preview
+schedules and the length of time until graduation.  This will lead to better
+decision making for what majors to pursue, and given the student's current
+state, which schedules are the most ideal in terms of saving time and resources.
+This will also help students save money, as they will be in school for fewer
+quarters. The university also benefits from the reduced labor cost and increased
+efficiency of the advising offices.
 
-\section{Problem Description} We focus on the classes offered within the Math
-department, although our algorithm and model can be easily extended to include
-over departments.
+\section{Problem Description} To make our examples concrete we will focus on the
+classes offered within the Math department, although our algorithm and model can
+be extended to include over departments.
 
-Every quarter the Math department offers a subset of classes from all of the
-available courses, each class is offered during a certain subset of available
-time periods. Since there are far more classes than time periods during the
-regular five-day week there are inevitably time conflicts between courses that
-a student could take. To graduate with a B.S. in Mathematics there are certain
-classes that are required and then there are elective requirements too. Adding
-to the problem is many classes cannot just be taken whenever they are available;
-instead, certain prerequisites must be satisfied first. In the case of math
-courses, there can be multiple different prerequisite sets that satisfy the
-requirements for a class.
+Every quarter the Math department offers a subset of courses from all of the
+available courses in the department. Each course is offered during a certain set
+time periods. There are inevitably time conflicts between courses that a student
+could take because there are far more courses offered than time periods during
+the regular five-day week  To graduate with a B.S. in Mathematics there are
+certain classes that are required and then there are elective requirements too.
+Additionally many classes have prerequisites which much be satisfied before they
+can be taken. In the case of math courses, there can be multiple, exclusive
+prerequisite sets that independently satisfy the requirements for a class.
 
@@
 \begin{itemize} \item A {\it course} refers to any course offered by the
-University of Washington. A course has associated with it a name used in the
-University of Washington time schedule, for example {\it MATH309}. A course also
-has associated with it the quarters in which it is offered. Throughout this
-paper, we will refer to these using their time schedule abbreviates: {\it F} for
-Fall Quarter, {\it W} for Winter Quarter, {\it Sp} for Spring Quarter, and {\it
-S} for Summer Quarter. Each course is also worth a specific amount of credits.
-\item We refer to a specific instance of a course as a {\it course offering}.
-Each course offering additionally has associated with it the time of the day at
-which the course meets, the quarter it is being offered, and the section number
-of the offering.  \end{itemize}
+University of Washington. A course is identified by its department name and
+number, for example {\it MATH309}. Each course is also worth a specific amount
+of credits. \item We refer to a specific instance of a course as a {\it course
+offering}. Each course offering additionally has associated with it the time of
+the day at which the course meets, the quarter it is being offered, and the
+section number of the offering. Throughout this paper, we will refer to quarters
+by their time schedule abbreviates: {\it F} for Fall Quarter, {\it W} for Winter
+Quarter, {\it Sp} for Spring Quarter, and {\it S} for Summer Quarter.
+\end{itemize}
 
-Additionally: \begin{itemize} \item Let $S$ be the set of courses required to be
-taken in order to graduate. $S$ can be thought of as our goal state, and can
-also contain elective courses a student simply wishes to take.  \item Let $T$ be
-the set of courses a given student has already completed. If no courses have
-been taken, this set may be empty.  \item Let $U$ be the set of all courses
-a student could possibly ever take.  \item Let $P$ be a set of rules specifying
-course prerequisites.  \item Let $l$ be the maximum number of credits a student
-wishes to take during any given quarter.  \item Let $O$ be the set of all
-50-minute time blocks between 8:30am and 7:30pm from Monday through Friday.
-\end{itemize}
+Additionally: \begin{itemize} \item Let $S$ be the set of courses required to
+graduate. $S$ can be thought of as our goal state, and can also contain elective
+courses a student simply wishes to take.  \item Let $T$ be the set of courses
+a given student has already completed. If no courses have been taken, this set
+may be empty.  \item Let $U$ be the set of all courses a student could possibly
+ever take.  \item Let $R$ be a set of rules specifying course prerequisites.
+\item Let $l$ be the maximum number of credits a student wishes to take during
+any given quarter.  \item Let $O$ be the set of all 50-minute time blocks
+between 8:30am and 7:30pm from Monday through Friday. \item Let $L_i$ be the set
+of courses a student can take in quarter i. \end{itemize}
 
@@
 existed between nodes $i$ and $j$ if and only if $i$ was a prerequisite for $j$.
-However, it does not work in our model, and courses can have a complicated set
-of multiple prerequisites.  For example, {\it MATH327} requires both $MATH126$
-and $MATH300$, or just $MATH136$.
+The DAG relationship works well when courses have single requirements. It
+becomes difficult, however, to track whether requires are fulfilled with courses
+that have multiple prerequisites, or worse courses that have alternative
+prerequisites. We decided that it was would be feasible to model prerequisites
+using a graph.
+\begin{figure} [ht] \begin{center}
+\includegraphics[scale=0.4]{prereq_logic_example} \end{center}
+\caption{Evaluation of a rule} \label{logic_ex} \end{figure}
 
-\begin{figure} [ht] 
-\begin{center} 
-\includegraphics[scale=0.4]{prereq_logic_example} 
-\end{center} 
-\caption{Evaluation of a rule} 
-\label{logic_ex} \end{figure}
+To represent alternative and multiple prerequisites we introduced a set of logic
+rules for each course. The prerequisites for each course are expressed as
+propositional logic statements stored in the set $R$. The literal propositions
+in each rule are related to whether or not a student has taken a particular
+course. For example, the rule associated with $MATH327$ in fig~\ref{logic_ex} is
+$$ \text{{\it MATH327: (MATH126 AND MATH300) OR MATH136}}.$$
 
-To solve this issues, we introduce a set of rules for each course. Each rule is
-a boolean expression represented in propositional logic in disjunctive normal
-form. For example, the rule associated with $MATH327$ in fig~\ref{logic_ex} is
-$$ \text{{\it MATH327: (MATH126 AND MATH300) OR MATH136}}.  $$ As previously
-stated, we denote the set of all rules for all classes as $P$.
-
-Furthermore, to determine if a given student is eligible to register for
-a course offering, we can evaluate the rule for the course, where each clause
-takes on the value true if a student has already taken the course, and false if
-he or she hasn't. If the rule evaluates to true, then the student is eligible to
-take the course; otherwise, the student is note. For example, consider
-figure~\ref{logic_ex}, here a student has previously taken courses $MATH126$ and
-$MATH300$. Since $MATH126$ and $MATH300$ have been taken, then those literals in
-the rule for $MATH327$ evaluate to true and therefore the entire rule is true.
-Since the entire rule is true, the student can take $MATH327$ when next
-available.
+To determine if a given student is eligible to register for a course offering,
+c we can evaluate the rule for c. Each literal proposition in the rule takes on
+the value true if a student has already taken the course, and false if he or she
+hasn't.  If the rule evaluates to true, then the student is eligible to take the
+course; otherwise. For example, consider Figure~\ref{logic_ex}, here a student
+has previously taken courses $MATH126$ and $MATH300$. Since $MATH126$ and
+$MATH300$ have been taken, then those literals in the rule for $MATH327$
+evaluate to true and therefore the entire rule is true.  Since the entire rule
+is true, the student can take $MATH327$.
 
 \subsection{Initial Model and Tractability} With a static schedule it would be
-possible to deterministically plan an entire schedule from start to finish,
-unfortunately we have neither the computational resources nor the luxury of
-a static schedule to do plan. From the student’s point of view, although there
-are some classes that are offered every quarter, senior electives are not, and
-sometimes it cannot be predicted more than one quarter in advance when a senior
-elective will be offered. Even given perfect information, where we know the
-possible class schedules for every quarter we face another problem; finding the
-best possible schedule reduces to enumerating possible class schedules for every
-quarter until the graduation requirements are satisfied. Given that we cannot
-predict in advance the conflicts between classes, we cannot even determine the
-branching factor of the enumeration tree nor its depth. Since we can only
-determine one quarter in advance, this suggests approaching the problem by using
-heuristics to pick the best possible classes each quarter.
+possible to plan an entire set of quarter by quarter schedules from start to
+finish. Unfortunately we have neither the computational resources nor the luxury
+of a static schedule. Some classes are offered every quarter others such as
+senior electives are not. Furthermore for some classes it cannot be predicted
+more than one quarter in advance when a senior elective will be offered. Even
+given perfect information, where we know the possible class schedules for every
+quarter we face another problem; finding the best possible schedule reduces to
+enumerating possible class schedules for every quarter until the graduation
+requirements are satisfied.  Given that we cannot predict in advance the
+conflicts between classes, we cannot even determine the branching factor of the
+enumeration tree nor its depth. Since we can only determine one quarter in
+advance, this suggests approaching the problem by using heuristics to pick the
+best possible classes each quarter.
 
-Instead of trying to solve the complete problem we chose to use the rolling
-horizon approach used by Xiangtong et al \cite{xiangton:informs} when scheduling
-classes for airlines. Our problem already has a well defined horizon
-demarcation, the quarters. For the rest of the paper we focus on developing
-methods to optimally pick classes in each quarter separately.
+Instead of trying to solve the complete problem we chose to use the "rolling
+horizon approach" used by Xiangtong et al \cite{xiangton:informs} when
+scheduling classes for airline pilots. Xiangtong et al. also could not develop
+an optimal class schedule in one pass, instead they picked optimal classes based
+using a heuristic over a well demarcated time period. Once they had a solution
+to the smaller time period, they increase the size of the time period, or rolled
+the horizon, and used the previously optimal solution to help determine
+a solution to the new larger problem. The team's algorithm continued rolling the
+time horizon on the problem until they had a solution to the entire problem.
 
-\section{Rolling Horizons Approach} We adapt a rolling horizons approach which
-uses a heuristic to pick the best possible schedule for any given quarter, given
-the classes offered during the quarter and the courses the student has
-previously taken. We use a greedy algorithm that functions on a per-quarter
-basis to accomplish this. The greedy algorithm greatly improves the tractability
-of course scheduling over an initially unknown number of quarters.
+\section{Picking Classes Per Quarter} Since determining an entire schedule at
+once is intractable we adapted the rolling horizons approach which uses
+a heuristic to pick the best possible schedule for a quarter. With a good
+heuristic we will be able to chose classes that lead to a minimized graduation
+time. Since we cannot analytically determine graduation time based on classes
+chosen in a single quarter, our heuristic tries to give high value to classes
+that would delay graduation is taken later rather than sooner.
 
@@
 classes we can currently take and then build conflict graphs for classes based
-on the times they are offered. In this section, we let $L$ be the initially
-empty set of course offerings a given student is eligible to take in the given
-quarter, $q$.
+on the times they are offered. The set $L_i$ is built in this part of the
+algorithm.
 
 \subsubsection{Procedure} We present the procedure for populating the set $L$ as
-follows: \begin{enumerate} \item For each rule in $P$ we iterate over each
-proposition in the rule, $p$, and check if $p \in T$. If $p \in T$ then we say
-$p$ evaluates to true; otherwise, $p$ evaluates to false. We recall that each
-rule is associated with a given course. If the entire rule evaluates to true,
-then we add the course associated with it to $L$. Once this is complete, $L$
-will consist of every course that could be taken based solely on prerequisites
-that we have fulfilled.  \item We now prune this set by removing all courses
-that are not offered in the current quarter. $L$ will necessarily now only
-contain classes that we can take given the current quarter and the classes we
-have already taken.  \item We now construct a conflict graph, $G_{conflict}$,
-where every course offering in $L$ is a node, and there is an edge between two
-nodes $i$ and $j$ if and only if the time of course offering $i$ overlaps the
-time of course offering $j$.  \end{enumerate}
+follows: \begin{enumerate} \item For each rule $r \in R$ we iterate over each
+proposition $p$ in the rule. Recall that $p$ represents a course, for example
+$MATH125$. We then check if $p \in T$.  If $p \in T$ then we say $p$ evaluates
+to true; otherwise, $p$ evaluates to false.  Using rules of logic we evaluate
+$r$, if $r$ true, then we add the course associated with it to $L$.  Once this
+is complete, $L$ will consist of every course that could be taken based solely
+on prerequisites that we have fulfilled.  \item We now prune $L$ by removing all
+courses that are not offered in the current quarter. $L$ now only contain
+classes that we can take given the current quarter and the classes we have
+already taken.  \item We now construct a conflict graph, $G_{conflict}$, where
+every course offering in $L$ is a node, and there is an edge between two nodes
+$i$ and $j$ if and only if the time of course offering $i$ overlaps the time of
+course offering $j$.  \end{enumerate}
+
+\begin{figure} [ht]
+    \begin{center}
+        \includegraphics[scale=0.35]{conflict_graph_ex}
+    \end{center}
+    \caption{A Conflict Graph}
+    \label{conflictg}
+\end{figure}
 
@@
 rule has twelve propositions, so we can bound the rule check time by a constant,
-$C$, and therefore step one can be done in $O(Cn)$ time.
+$C$, and therefore step one can be done in $O(n)$ time.
 
-Continuing to step two, the worst case is when $L$ = $Q$. Checking whether
-a course is offered in a given quarter can be done in constant time, and we have
-$n$ courses to check, therefore step two can also be done in $O(Cn)$ time.
+Continuing to step two, the worst case is when $L$ = {all classes offered during
+the quarter}.  Checking whether a course is offered in a given quarter can be
+done in constant time, and we have $n$ courses to check, therefore step two can
+also be done in $O(n)$ time.
 
@@
 five days of the week, and therefore we have 45 time slots to check in our
-conflict graph, again yielding $O(Cn)$ runtime.
+conflict graph, again yielding $O(n)$ runtime.
 
@@
 
-\subsection{Selection of Quarterly Classes} \subsubsection{Selection Procedure}
-At this point in time, we know which course offerings are available for the
-given student to register for, and we must now pick the most optimal ones.  We
-present two ways of approaching quarterly class choice. We will initially encode
-the class selection as a binary integer programming problem in spirit of the
-methods presented by Dinkel et al \cite{dinkel:scheduling}. and Pritsker et al
-\cite{prisker:informs}. A second approach developed by Yamada et al
-\cite{yamada:heuristic} and Pferschy et al \cite{pferschy:kcg} is presented for
-dealing with very large class choices where binary LP would be difficult solve.
+\subsection{Selection of Quarterly Classes} To select classes we give each class
+a value and weight. The primary problem is the the classes contained in $L$
+almost certainly have time conflicts, likely between courses that need to be
+taken.
 
@@
 should be picked in a given schedule, we use a heuristic function which assigns
-each course offering a value. The higher the value of this function for a given
-course offering is, the more likely the given course offering will be placed in
-the schedule. Recalling that our goal is to minimize the total number of
-quarters taken (and implicitly, maximize the total number of courses taken per
-quarter). Our heuristic function should give higher value to course offerings
-that will help us achieve this goal. One component to the heuristic function
-then must be the number of dependent courses a given course offering has. If
-a course satisfies many prerequisites, then we know that there are many courses
-that we cannot complete without it, so we want our heuristic function to favor
-it.
+each class a value. The higher the value of this function for a class, the more
+likely that class will be placed in the schedule.  Recalling that our goal is to
+minimize the total number of quarters taken (and implicitly, maximize the total
+number of courses taken per quarter).  Our heuristic function should give higher
+value to class that maximize the number of classes taken this quarter and in
+future quarters.  One component to the heuristic function then must be the
+number of dependent courses a given course offering has. If a course satisfies
+many prerequisites, then we know that there are many courses that we cannot
+complete without it, so we want our heuristic function to favor it.
 
@@
 
-Where $c$ is a given course offering,
-$q$ is the current quarter, and $\alpha$, $\beta$, and $\gamma$ are real
-numbers, giving a weight to piece of the heuristic. The scalars are determined
-experimentally.
+Where $c$ is a given course offering, $q$ is the current quarter, and $\alpha$,
+$\beta$, and $\gamma$ are real numbers, giving a weight to piece of the
+heuristic. The scalars are determined experimentally.
+
+\subsubsection{Selection Procedure} At this point in the algorithm we know which
+courses are available to a student and their relative values. We present two
+ways of approaching quarterly class choice.  We will initially encode the class
+selection as a binary integer programming problem in spirit of the methods
+presented by Dinkel et al \cite{dinkel:scheduling}. and Pritsker et al
+\cite{prisker:informs}. A second approach developed by Yamada et al
+\cite{yamada:heuristic} and Pferschy et al \cite{pferschy:kcg} is presented for
+dealing with very large class choices where binary LP would be difficult solve.
 
@@
 We now state our problem as follows:
+
 Maximize: 
-
 \begin{equation}
@@
 \begin{equation}
-    \forall t \in O,X_i+X_j \leq 1, \forall pairs (X_i,X_j) \in t
+    \forall t \in O,X_i+X_j \leq 1, \forall \text{ connected pairs } (X_i,X_j) \in t
     \label{time_con}
+\end{equation}
+
+\begin{equation}
+   X_i + X_j \leq 1, \forall \text{ pairs } (X_i,X_j) \text{ where } X_i \text{and} X_j \text{ are offerings of
+   the same course}
+    \label{course_con}
 \end{equation}
@@
 
-As noted in the reference literature this is actually a (strongly) NP-complete
-problem as described \cite{pferschy:kcg}. For the rest of this section we
-consider an alternative formulation of class selection as a knapsack problem
-with conflict graphs and suggest the use a technique developed by Yamada et. al.
-\cite{yamada:heuristic}.
-
-We begin by simplifying $G_{conflict}$ to be a fully connected graph by adding
-dummy nodes $v'$ and insert edges that connect $v'$ to every other node in the
-graph. To ensure that $v'$ does not alter the final optimal solution, we assign
-it the following properties: \begin{itemize} \item $Weight(v') = l+1$ \item
-$Value(v') = 0$ \end{itemize} As the weight of our dummy node is greater than
-our credit limit, and it's value is zero, we know that it will never appear in
-the optimal solution. With the above reasoning we show for our case of class
-schedule that Pferschy and Schauer's \cite{pferschy:kcg} claim that adding
-a dummy node without loss of generality holds. We can now conveniently write the
-class picking problem as the knapsack with conflict graphs problem (KCG):
-
-Maximize:
-
-\begin{equation}
-    \text{Value of Quarter} = \sum_{i=1}^m Value(c_i) * X_i
-\end{equation}
-
-Subject To:
-
-\begin{equation}
-\sum_{i=1}^m Weight(c_i) * X_i \leq l,
-\end{equation}
-
-\begin{equation}
-X_i + X_j \leq 1 \text{ for all connected vertices } i, j \in G_{conflict},
-\end{equation}
-
-\begin{equation}
-X_i \in \{0, 1\}
-\end{equation}
-
-For versions of this problem in the low hundreds of variables it is possible 
-to solve this problem exactly using modern LP solving software 
-\cite{yamada:heuristic}. In our case, just solving for the optimal classes for 
-a single major fits well into the scope of a few hundred (or even just tens of) 
-variables. For multiple majors or to consider class selection
+As noted in the reference literature this is actually a NP-hard
+problem as described \cite{pferschy:kcg}. For versions of this problem in the
+low hundreds of variables, however it is still possible to solve using modern LP
+solving software \cite{yamada:heuristic}. In our case, just solving for the
+optimal classes for a single major fits well into the scope of a few hundred (or
+even just tens of) variables. For multiple majors or to consider class selection
 for distribution and elective requirements, however the number of classes
 balloons significantly and the KCG problem becomes much harder. For this version
-of the problem we suggest using the techniques used by Yamada et al
+of the problem we suggest the techniques used by Yamada et al
 \cite{yamada:heuristic}, which we briefly outline below.
+
+For the rest of this section we consider an alternative formulation of class
+selection as a knapsack problem with conflict graphs and suggest the use
+a technique developed by Yamada et. al.  \cite{yamada:heuristic}.
 

### 2010-12-11 quarter_page_extracter.py
added commet

@@ -45,6 +45,7 @@ def main(args):
         if math_class_match is not None:
             print("Matched "+math_class_match.group(1))
             curr_class = int(math_class_match.group(1))
+        # next pri is switching the offered sections of a course
         elif main_section_match is not None:
             curr_sec = offering(main_section_match.group(1),
                 main_section_match.group(2),main_section_match.group(3),
@@ -54,6 +55,7 @@ def main(args):
                 classes[curr_class].update({curr_sec:[]})
             except KeyError:
                 classes[curr_class] = {curr_sec:[]}
+        # finally we consider the qz sections offered
         elif qz_section_match is not None:
             curr_qz = offering(qz_section_match.group(1),
                 qz_section_match.group(2),0,

### 2010-12-11 paper/paper.tex
Rewrite/edit sections 1,2....again.

@@
 
-\abstract{The Bachelor's Degree is designed to take four years to complete,
-however, many students fail to complete it this amount of time. One 
+\abstract{The Bachelor's Degree is designed to take four years to complete;
+however, many students fail to complete it in this time frame. One 
 factor contributing to this is poor planning when registering for classes.
-We present an algorithm to assist students in
-constructing a schedule that works towards completing a degree in the shortest
-amount of time possible.}
+We present an algorithm to assist students in constructing a schedule 
+that works towards completing a degree in the shortest amount of time 
+possible. The algorithm presented uses heurisitics to attempt to pick the best
+class set each quarter, working towards a complete schedule quartery by 
+quarter.}
 \pagebreak
@@
 \section{Introduction} \subsection{Background} Registration can be a stressful 
-time for every student at the University of Washington. Without good
-planning, students can make poor registration choices resulting in more quarter
-being needed to graduate than otherwise necessary. Picking the correct class
-involves considering quarterly class offerings, class credits, the prerequisites
-of a class, time conflicts, and how to fulfill future prerequisites. The
-University of Washington registration system already provides a tool to aid
-students, schedule finder. Schedule finder however, only takes into account time
-conflicts in one quarter, and does not consider what classes are best for
-minimizing graduation time.
-
-A student's graduation time may be extended because of his or her inability to
-take a class that has limited offerings during that particular year.
+time for students at the University of Washington. Without proper planning, 
+a student can run the risk of making poor registration choices, which can 
+result in the student taking more quarters than otherwise necessary. However, 
+choosing the right schedule is complex and there are multiple variables to 
+consider. Picking classes involves the consideration of quarterly class 
+offerings, class credits, prerequisites, time conflicts, and whether or not 
+classes will fulfill future prerequisites. All of these factors can affect a 
+student’s graduation time, especially if that class has limited offerings 
+during a particular year. The University of Washington registration system 
+already provides a tool to aid students: schedule finder. However, schedule 
+finder only takes into account time conflicts per quarter, and not what 
+classes are best for minimizing graduation time.
 
 \subsection{Goal} Our goal is to provide an automated system that gives quarter
-by quarter schedules until a student graduates. As input our system will take in
+by quarter schedules up until graduates. As input our system will take in
 the current set of classes a student has taken and the preferred credit hours
@@
 \section{Problem Description} To make our examples concrete we will focus on the
-classes offered within the Math department, although our algorithm and model can
-be extended to include over departments.
+classes offered within the Math department; however, our algorithm and model can
+be extended to include other departments.
 
-Every quarter the Math department offers a subset of courses from all of the
-available courses in the department. Each course is offered during a certain set
-time periods. There are inevitably time conflicts between courses that a student
+Every quarter the Mathematics department offers a selection of courses; not every
+course is offered every quarter. Each course is offered during a certain set
+time periods. There are inevitable time conflicts among courses that a student
 could take because there are far more courses offered than time periods during
-the regular five-day week  To graduate with a B.S. in Mathematics there are
-certain classes that are required and then there are elective requirements too.
+the regular five-day week. Graduating with a B.S. in Mathematics requires
+specific classes in addition to elective requirements.
 Additionally many classes have prerequisites which much be satisfied before they
@@
 of credits. \item We refer to a specific instance of a course as a {\it course
-offering}. Each course offering additionally has associated with it the time of
+offering}. Additionally, each course offering  has associated with it the time of
 the day at which the course meets, the quarter it is being offered, and the
@@
 The DAG relationship works well when courses have single requirements. It
-becomes difficult, however, to track whether requires are fulfilled with courses
+becomes difficult, however, to track whether requirements are fulfilled with courses
 that have multiple prerequisites, or worse courses that have alternative
-prerequisites. We decided that it was would be feasible to model prerequisites
+prerequisites. We decided that it would be feasible to model prerequisites
 using a graph.
@@
 a heuristic to pick the best possible schedule for a quarter. With a good
-heuristic we will be able to chose classes that lead to a minimized graduation
+heuristic we will be able to choose classes that lead to a minimized graduation
 time. Since we cannot analytically determine graduation time based on classes

### 2010-12-11 paper/paper.tex
added new section describing quiz section handling and two new graphics

--- 2010-12-11 paper/paper.tex
@@
 
-\section{Problem Description} To make our examples concrete we will focus on the
-classes offered within the Math department, although our algorithm and model can
-be extended to include over departments.
+\section{Problem Description} \label{pdesc} To make our examples concrete we
+will focus on the classes offered within the Math department, although our
+algorithm and model can be extended to include over departments.
 
@@
 
+\section{Model Elements} The solve the problem outlined in section \ref{pdesc}
+we formalized elements of the academic quarter. Some of the hardest elements to
+model were representing prerequisites for a course, how to handle quiz sections,
+and simply choosing appropriate sets to contain information about a student's
+academic progress.
+
+\subsection{Classes and Quiz Sections} Another modeling difficulty that we
+initially ignore was how to represent quiz sections for a class. Each time
+a course is offered is may be offered as multiple classes. For instance, during
+$F$ $MATH307$ is offered as 7 different sections all teaching the same material
+during different time periods. Within a specific quarter schedule line number,
+SLNs uniquely identify courses allowing us to differentiate between classes with
+the same name. Continuing the above example, the SLNs of the 7 $MATH307$
+instances may be 15432, 15433, 15434...
+
+Our peer review team commented that we had not accounted for courses that
+required concurrent quiz sections. First we note that the quiz sections for
+a class never occur during the same time periods as the class itself (otherwise
+students would not be able to attend the quiz section). Quiz sections also have
+unique SLNs, to include a required quiz section we include the quiz section SLN
+with the SLN for the associated class. Finally, courses such at $MATH125$ can
+have multiple sections, each with several quiz sections of which only one need
+be taken. We found the easiest way to represent this situation was to iterate
+over the quiz sections for each class section and generate a puesdo class for
+each pair of class section and quiz section that occured during the time periods
+for both quiz and main sections. Fig. \ref{quizzes} contains one section
+$MATH125$ being split across 3 quiz sections.
+\begin{figure} [ht] 
+    \begin{center}
+        \includegraphics[scale=0.4]{quiz_sections} 
+    \end{center}
+    \caption{Splitting up MATH125} 
+    \label{quizzes} 
+\end{figure}
+
 \subsection{Definitions} We will now formalize some terms used in our paper.
@@
 number, for example {\it MATH309}. Each course is also worth a specific amount
-of credits. \item We refer to a specific instance of a course as a {\it course
-offering}. Each course offering additionally has associated with it the time of
-the day at which the course meets, the quarter it is being offered, and the
-section number of the offering. Throughout this paper, we will refer to quarters
-by their time schedule abbreviates: {\it F} for Fall Quarter, {\it W} for Winter
-Quarter, {\it Sp} for Spring Quarter, and {\it S} for Summer Quarter.
-\end{itemize}
+of credits. \item We refer to a specific instance of a course as a {\it class}.
+Each class is has associated with it the time of the day at which it  meets, the
+quarter it is being offered, and the section number of the offering.  Throughout
+this paper, we will refer to quarters by their time schedule abbreviates: {\it
+F} for Fall Quarter, {\it W} for Winter Quarter, {\it Sp} for Spring Quarter,
+and {\it S} for Summer Quarter. \end{itemize}
 
@@
 using a graph.
-\begin{figure} [ht] \begin{center}
-\includegraphics[scale=0.4]{prereq_logic_example} \end{center}
-\caption{Evaluation of a rule} \label{logic_ex} \end{figure}
+\begin{figure} [ht] 
+    \begin{center}
+        \includegraphics[scale=0.4]{prereq_logic_example} 
+    \end{center}
+    \caption{Evaluation of a rule} 
+    \label{logic_ex} 
+\end{figure}
 
@@
 
-\subsection{Initial Model and Tractability} With a static schedule it would be
+\section{Tractability} With a static schedule it would be
 possible to plan an entire set of quarter by quarter schedules from start to
@@
 $i$ and $j$ if and only if the time of course offering $i$ overlaps the time of
-course offering $j$.  \end{enumerate}
-
+course offering $j$. An example conflict graph for one time period is given in
+Fig. \ref{conflictg}. \end{enumerate} 
 \begin{figure} [ht] 

### 2010-12-11 paper/paper.tex

--- 2010-12-11 paper/paper.tex
@@
 
-\section{Problem Description} To make our examples concrete we will focus on the
+\section{Problem Description} \label{psdesc} To make our examples concrete we will focus on the
 classes offered within the Math department; however, our algorithm and model can
@@
 number, for example {\it MATH309}. Each course is also worth a specific amount
-of credits. \item We refer to a specific instance of a course as a {\it course
-offering}. Additionally, each course offering  has associated with it the time of
-the day at which the course meets, the quarter it is being offered, and the
-section number of the offering. Throughout this paper, we will refer to quarters
-by their time schedule abbreviates: {\it F} for Fall Quarter, {\it W} for Winter
-Quarter, {\it Sp} for Spring Quarter, and {\it S} for Summer Quarter.
-\end{itemize}
+of credits. \item We refer to a specific instance of a course as a {\it class}.
+Each class has associated with it the time of the day at which it meets, the
+quarter it is being offered, and the section number of the offering.  Throughout
+this paper, we will refer to quarters by their time schedule abbreviates: {\it
+F} for Fall Quarter, {\it W} for Winter Quarter, {\it Sp} for Spring Quarter,
+and {\it S} for Summer Quarter. \end{itemize}
 

### 2010-12-11 paper/paper.tex
Added equation references, section references in section 5

@@
 that works towards completing a degree in the shortest amount of time 
-possible. The algorithm presented uses heurisitics to attempt to pick the best
-class set each quarter, working towards a complete schedule quartery by 
+possible. The algorithm presented uses heuristics to attempt to pick the best
+class set each quarter, working towards a complete schedule quarterly by 
 quarter.}
@@
 
-\section{Problem Description} \label{psdesc} To make our examples concrete we will focus on the
+\section{Problem Description} \label{pdesc} To make our examples concrete we will focus on the
 classes offered within the Math department; however, our algorithm and model can
@@
 
-\subsection{Classes and Quiz Sections} Another modeling difficulty that we
-initially ignore was how to represent quiz sections for a class. Each time
-a course is offered is may be offered as multiple classes. For instance, during
-$F$ $MATH307$ is offered as 7 different sections all teaching the same material
-during different time periods. Within a specific quarter schedule line number,
-SLNs uniquely identify courses allowing us to differentiate between classes with
-the same name. Continuing the above example, the SLNs of the 7 $MATH307$
-instances may be 15432, 15433, 15434...
+\subsection{Classes and Quiz Sections} \label{classDefs} Another modeling
+difficulty that we initially ignore was how to represent quiz sections for
+a class. Each time a course is offered is may be offered as multiple classes.
+For instance, during $F$ $MATH307$ is offered as 7 different sections all
+teaching the same material during different time periods. Within a specific
+quarter schedule line number, SLNs uniquely identify courses allowing us to
+differentiate between classes with the same name. Continuing the above example,
+the SLNs of the 7 $MATH307$ instances may be 15432, 15433, 15434...
 
@@
 be taken. We found the easiest way to represent this situation was to iterate
-over the quiz sections for each class section and generate a puesdo class for
-each pair of class section and quiz section that occured during the time periods
+over the quiz sections for each class section and generate a pseudo class for
+each pair of class section and quiz section that occurred during the time periods
 for both quiz and main sections. Fig. \ref{quizzes} contains one section
@@
 won't have the opportunity to take it for another entire year.
-
 \begin{figure} [ht]
@@
 \cite{yamada:heuristic} and Pferschy et al \cite{pferschy:kcg} is presented for
-dealing with very large class choices where binary LP would be difficult solve.
+dealing with very large class choices where a binary LP would be difficult
+solve.
 
@@
     \text{Value of Quarter} = \sum_{i=1}^m Value(c_i) * X_i
-    \label{qtr_val}
+    \label{qtrval}
 \end{equation}
@@
     \sum_{i=1}^m Weight(c_i) * X_i \leq l, 
-    \label{weight_lim}
+    \label{weightlim}
 \end{equation}
@@
     \forall t \in O,X_i+X_j \leq 1, \forall \text{ connected pairs } (X_i,X_j) \in t
-    \label{time_con}
+    \label{timecon}
 \end{equation}
@@
    the same course}
-    \label{course_con}
+    \label{coursecon}
 \end{equation}
 
-The linear program
-can be interpreted as follows: \begin{itemize} \item The value of a quarter is
-the total number of courses taken that quarter.  \item In each quarter, the
-total number of credits cannot exceed the specified limit.  \item Only one
-course can be taken in one particular time slot \item A course can only be taken
-or not taken. In other words, decision variables are binary.  \item Each
-non-zero decision variable in the optimal solution represents a class we will
-take.  \end{itemize}
+The linear program can be interpreted as follows: \begin{itemize} \item
+Eqn.~\ref{qtrval}: The value of a quarter is the total number of courses taken
+that quarter. We want to maximize this value, to maximize the number of high
+value classes we take each quarter. \item Eqn.~\ref{weightlim}: In each quarter,
+the total number of credits cannot exceed the specified limit. $Weight(c_i)$
+represents a short hand way of indicating the number of credits hours a class
+takes. \item Eqn.~\ref{timecon}: Only one course can be taken in one particular
+time slot, this is where $G_{conflict}$ is used. Without the prebuilt conflict
+graph it would be difficult to build the set of constraints based on equation
+number~ref{timecon}. \item Eqn.~\ref{coursecon}: Previously in
+section~\ref{classDefs} we noted that in many cases a course is offered in
+multiple sections during the same quarter. A student can only take one offering
+of a course, this constraint eqn must be built for every class-quiz section
+combination which represent that same course.\end{itemize}
 
-Each of the equations based on eqn.~\ref{time_con} ensure that we only take one class per 
-time period per day. Also this linear program is ‘nice’ in that it always has 
-feasible origin (don’t take any classes), and that the simplest formulation 
-is already in standard form.
+Solving this LP will yield a set $X_i$s that are 1. We represent "taking" these
+classes by moving them into the set $T$. Then we advance the current quarter,
+rolling the horizon back and run the algorithm again, repeating as necessary
+till we have a complete schedule.
 
-As noted in the reference literature this is actually a NP-hard
-problem as described \cite{pferschy:kcg}. For versions of this problem in the
-low hundreds of variables, however it is still possible to solve using modern LP
-solving software \cite{yamada:heuristic}. In our case, just solving for the
-optimal classes for a single major fits well into the scope of a few hundred (or
-even just tens of) variables. For multiple majors or to consider class selection
-for distribution and elective requirements, however the number of classes
-balloons significantly and the KCG problem becomes much harder. For this version
-of the problem we suggest the techniques used by Yamada et al
-\cite{yamada:heuristic}, which we briefly outline below.
+\subsection{Alternative Formulation} As noted in the reference literature this
+is actually a NP-hard problem as described \cite{pferschy:kcg}.  For versions of
+this problem in the low hundreds of variables, however it is still possible to
+solve using modern LP solving software \cite{yamada:heuristic}. In our case,
+just solving for the optimal classes for a single major fits well into the scope
+of a few hundred (or even just tens of) variables. For multiple majors or to
+consider class selection for distribution and elective requirements, however the
+number of classes balloons significantly and the KCG problem becomes much
+harder. For this version of the problem we suggest the techniques used by Yamada
+et al \cite{yamada:heuristic}, which we briefly outline below.
 

### 2010-12-12 paper/paper.tex
elminate some passive voice and weasel words; additional editing / improvements

@@
 \usepackage{array} % for better arrays (eg matrices) in maths
-\usepackage{paralist} % very flexible & customisable lists (eg. enumerate/itemize, etc.)
+\usepackage{paralist} % flexible & customisable lists (eg. enumerate/itemize, etc.)
 \usepackage{epsfig}
@@
 %%% HEADERS & FOOTERS
-\usepackage{fancyhdr} % This should be set AFTER setting up the page geometry
+\usepackage{fancyhdr} % Set AFTER setting up the page geometry
 \pagestyle{fancy} % options: empty , plain , fancy
@@
 \author{Ruyi Cai \and Matthew Fonda \and Michael Hughes}
-%\date{} % Activate to display a given date or no date (if empty),
-         % otherwise the current date is printed
 
@@
 
-\abstract{The Bachelor's Degree is designed to take four years to complete;
-however, many students fail to complete it in this time frame. One 
+\abstract{The Bachelor's Degree takes four years to complete;
+however, students often fail to complete it in this time frame. One
 factor contributing to this is poor planning when registering for classes.
@@
 time for students at the University of Washington. Without proper planning, 
-a student can run the risk of making poor registration choices, which can 
-result in the student taking more quarters than otherwise necessary. However, 
-choosing the right schedule is complex and there are multiple variables to 
+a student runs the risk of making poor registration choices resulting
+in the student taking more quarters than otherwise necessary.
+Choosing the right schedule is complex and there are multiple variables to
 consider. Picking classes involves the consideration of quarterly class 
 offerings, class credits, prerequisites, time conflicts, and whether or not 
-classes will fulfill future prerequisites. All of these factors can affect a 
-student’s graduation time, especially if that class has limited offerings 
-during a particular year. The University of Washington registration system 
-already provides a tool to aid students: schedule finder. However, schedule 
-finder only takes into account time conflicts per quarter, and not what 
-classes are best for minimizing graduation time.
+classes will fulfill future prerequisites. All of these factors affect the amount
+of time required for a student to graduate. The University of Washington registration
+system already provides a tool to aid students: schedule finder. However, schedule
+finder only takes into account time conflicts per quarter, and not what classes are
+best for minimizing graduation time.
 
-\subsection{Goal} Our goal is to provide an automated system that gives quarter
-by quarter schedules up until graduates. As input our system will take in
-the current set of classes a student has taken and the preferred credit hours
-per quarter. Additional input will be the current and past university time
-schedules. As output our system will provide a class schedule that minimizes the
-number of quarters until graduation.
+\subsection{Goal} Our goal is to provide an automated system that generates quarterly
+schedules until the point of graduation. As input our system will take in
+the current set of classes a student has taken, the preferred credit hours
+per quarter of a student, and current and past university time schedules.
+Our system will output a class schedule that minimizes the
+number of quarters required until graduation.
 
 \subsection{Importance} The implementation of our system as part of the school’s
-registration system will benefit the students by enabling them to map out their
+registration system will benefit students by enabling them to map out their
 quarterly class schedules until graduation. Students will be able to preview
@@
 quarters. The university also benefits from the reduced labor cost and increased
-efficiency of the advising offices.
+efficiency of advising offices.
 
 \section{Problem Description} \label{pdesc} To make our examples concrete we will focus on the
-classes offered within the Math department; however, our algorithm and model can
-be extended to include other departments.
+classes offered within the Math department; however, our algorithm also works with for other departments.
 
-Every quarter the Mathematics department offers a selection of courses; not every
-course is offered every quarter. Each course is offered during a certain set
+Every quarter the Mathematics department offers a selection of courses. This offering is a
+subset of all Mathematics courses, and varies from quarter to quarter.
+Each course is available within a specific set of
 time periods. There are inevitable time conflicts among courses that a student
-could take because there are far more courses offered than time periods during
+could take because there are far more courses offered than there are time periods during
 the regular five-day week. Graduating with a B.S. in Mathematics requires
 specific classes in addition to elective requirements.
-Additionally many classes have prerequisites which much be satisfied before they
-can be taken. In the case of math courses, there can be multiple, exclusive
-prerequisite sets that independently satisfy the requirements for a class.
+Additionally most courses require the completion of prerequisite courses.
+There can be multiple, disjoint prerequisite sets that independently
+satisfy the requirements for a class.
 
 Our goal is to determine a schedule of classes from some starting point to
-graduation that given the above properties minimizes the total number of
-quarters needed. Explicitly: \\ Minimize the total
+graduation that given the above properties, minimizes the total number of
+quarters required to graduate. Explicitly: \\ Minimize the total
 number of quarters needed to graduate subject to: \begin{enumerate} \item Only
@@
 
-\subsection{Classes and Quiz Sections} \label{classDefs} Another modeling
-difficulty that we initially ignore was how to represent quiz sections for
-a class. Each time a course is offered is may be offered as multiple classes.
-For instance, during $F$ $MATH307$ is offered as 7 different sections all
-teaching the same material during different time periods. Within a specific
-quarter schedule line number, SLNs uniquely identify courses allowing us to
-differentiate between classes with the same name. Continuing the above example,
-the SLNs of the 7 $MATH307$ instances may be 15432, 15433, 15434...
+\subsection{Classes and Quiz Sections} \label{classDefs} A modeling
+difficulty that we initially ignore was accounting for quiz sections and
+for multiple offering of the same course within one quarter.
+For instance, the Fall Quarter time schedule includes seven different
+offerings of $MATH307$. A schedule line number (SLN)
+uniquely identifies each class, allowing us to differentiate between classes
+with the same title. Continuing the above example,
+the SLNs of the seven $MATH307$ offerings may be 15432, 15433, 15434...
 
-Our peer review team commented that we had not accounted for courses that
+Our peer review team commented that we did not account for courses that
 required concurrent quiz sections. First we note that the quiz sections for
@@
 unique SLNs, to include a required quiz section we include the quiz section SLN
-with the SLN for the associated class. Finally, courses such at $MATH125$ can
-have multiple sections, each with several quiz sections of which only one need
-be taken. We found the easiest way to represent this situation was to iterate
-over the quiz sections for each class section and generate a pseudo class for
-each pair of class section and quiz section that occurred during the time periods
-for both quiz and main sections. Fig. \ref{quizzes} contains one section
-$MATH125$ being split across 3 quiz sections.
+with the SLN for the associated class. Finally, courses can have multiple sections
+and multiple quiz sections. Additionally, we have the restriction that students
+cannot concurrently enroll in more than one section of a given course. We found that
+the easiest way of representing this situation is generating a pseudo-class for each
+possible class and quiz section pair.
+Fig. \ref{quizzes} illustrates splitting a $MATH125$ section into three pseudo-classes.
 \begin{figure} [ht] 
@@
 
-\subsection{Definitions} We will now formalize some terms used in our paper.
+\subsection{Definitions} We now formalize some terms used in our paper.
 \begin{itemize} \item A {\it course} refers to any course offered by the
-University of Washington. A course is identified by its department name and
-number, for example {\it MATH309}. Each course is also worth a specific amount
+University of Washington. A course is uniquely identified by its department name and
+number, for example {\it MATH309}. Each course is worth a specific amount
 of credits. \item We refer to a specific instance of a course as a {\it class}.
-Each class has associated with it the time of the day at which it meets, the
-quarter it is being offered, and the section number of the offering.  Throughout
-this paper, we will refer to quarters by their time schedule abbreviates: {\it
+Each class meets at a specific time of day during a specific quarter. Additionally,
+each class has a unique SLN. Throughout
+this paper, we refer to quarters by their time schedule abbreviates: {\it
 F} for Fall Quarter, {\it W} for Winter Quarter, {\it Sp} for Spring Quarter,
@@
 
-Additionally: \begin{itemize} \item Let $S$ be the set of courses required to
-graduate. $S$ can be thought of as our goal state, and can also contain elective
-courses a student simply wishes to take.  \item Let $T$ be the set of courses
-a given student has already completed. If no courses have been taken, this set
-may be empty.  \item Let $U$ be the set of all courses a student could possibly
-ever take.  \item Let $R$ be a set of rules specifying course prerequisites.
+Additionally: \begin{itemize} \item Let $S$ be the set of all courses required for
+a student to graduate, additionally including any electives a student wishes to take.
+$S$ is our goal state.  \item Let $T$ be the set of courses
+a given student has already completed.
+\item Let $R$ be a set of rules specifying course prerequisites.
 \item Let $l$ be the maximum number of credits a student wishes to take during
@@
 
-\subsection{Representing Prerequisites} Initially, we hoped to represent course
-prerequisites using a directed acyclic graph. Each node was a course, and edges
+\subsection{Representing Prerequisites} Initially, we represented course
+prerequisites using a directed acyclic graph. Each node represented a course, and edges
 existed between nodes $i$ and $j$ if and only if $i$ was a prerequisite for $j$.
-The DAG relationship works well when courses have single requirements. It
-becomes difficult, however, to track whether requirements are fulfilled with courses
-that have multiple prerequisites, or worse courses that have alternative
-prerequisites. We decided that it would be feasible to model prerequisites
-using a graph.
+This model works well if each course has only one prerequisite. However,
+this model fails to represent courses that have multiple prerequisites
+or courses that have alternative prerequisites. This is the case for some courses.
+We therefore concluded that it is infeasible to model prerequisites using a graph.
 \begin{figure} [ht] 
@@
 To represent alternative and multiple prerequisites we introduced a set of logic
-rules for each course. The prerequisites for each course are expressed as
-propositional logic statements stored in the set $R$. The literal propositions
-in each rule are related to whether or not a student has taken a particular
-course. For example, the rule associated with $MATH327$ in fig~\ref{logic_ex} is
+rules for each course. The set $R$ consists of a prepositional logic statements representing
+the prerequisites of each course. The literals in each statement are a course.
+For example, the rule associated with $MATH327$ in Figure~\ref{logic_ex} is
 $$ \text{{\it MATH327: (MATH126 AND MATH300) OR MATH136}}.$$
 
-To determine if a given student is eligible to register for a course offering,
-c we can evaluate the rule for c. Each literal proposition in the rule takes on
-the value true if a student has already taken the course, and false if he or she
-hasn't.  If the rule evaluates to true, then the student is eligible to take the
-course; otherwise. For example, consider Figure~\ref{logic_ex}, here a student
-has previously taken courses $MATH126$ and $MATH300$. Since $MATH126$ and
-$MATH300$ have been taken, then those literals in the rule for $MATH327$
-evaluate to true and therefore the entire rule is true.  Since the entire rule
-is true, the student can take $MATH327$.
+To determine if a given student is eligible to register for a given class $c$,
+we evaluate the rule for $c$. Each literal in the statement takes on
+the value true if a student has already taken the course, and false if the student
+hasn't. If the statement evaluates to true, then the student is eligible to take the
+course; otherwise, the student is not. For example, consider Figure~\ref{logic_ex}. Here, a student
+has previously taken courses $MATH126$ and $MATH300$. Since the student completed both $MATH126$ and
+$MATH300$, these literals in the rule for $MATH327$ evaluate to true,
+and therefore the entire rule evaluates true. The student is therefore eligible to take $MATH327$.
 
@@
 finish. Unfortunately we have neither the computational resources nor the luxury
-of a static schedule. Some classes are offered every quarter others such as
-senior electives are not. Furthermore for some classes it cannot be predicted
-more than one quarter in advance when a senior elective will be offered. Even
-given perfect information, where we know the possible class schedules for every
+of a static schedule. Furthermore, the next offering of a given course is unpredictable.
+Even with perfect information, where we know the possible class schedules for every
 quarter we face another problem; finding the best possible schedule reduces to
-enumerating possible class schedules for every quarter until the graduation
-requirements are satisfied.  Given that we cannot predict in advance the
+enumerating possible class schedules for every quarter until the fulfillment of graduation
+requirements. Given that we cannot predict in advance the
 conflicts between classes, we cannot even determine the branching factor of the
@@
 scheduling classes for airline pilots. Xiangtong et al. also could not develop
-an optimal class schedule in one pass, instead they picked optimal classes based
+an optimal class schedule in one pass, instead they picked optimal classes
 using a heuristic over a well demarcated time period. Once they had a solution
-to the smaller time period, they increase the size of the time period, or rolled
+to the smaller time period, they increased the size of the time period, or rolled
 the horizon, and used the previously optimal solution to help determine
@@
 \section{Picking Classes Per Quarter} Since determining an entire schedule at
-once is intractable we adapted the rolling horizons approach which uses
+once is intractable, we adapted the rolling horizons approach which uses
 a heuristic to pick the best possible schedule for a quarter. With a good
@@
 time. Since we cannot analytically determine graduation time based on classes
-chosen in a single quarter, our heuristic tries to give high value to classes
-that would delay graduation is taken later rather than sooner.
+chosen in a single quarter, our heuristic gives higher value to classes
+that delay graduation the longest.
 
@@
 \subsection{Determining Available Course Offerings} We need to select the
-classes we can currently take and then build conflict graphs for classes based
-on the times they are offered. The set $L_i$ is built in this part of the
-algorithm.
+classes we can currently take and then build time conflict graphs.
+We construct the set $L_i$ during this phase of the algorithm.
 
@@
 follows: \begin{enumerate} \item For each rule $r \in R$ we iterate over each
-proposition $p$ in the rule. Recall that $p$ represents a course, for example
-$MATH125$. We then check if $p \in T$.  If $p \in T$ then we say $p$ evaluates
+literal $p$ in the rule. Recall that $p$ represents a course, for example
+$MATH125$. We then check if $p \in T$.  If $p \in T$, we say $p$ evaluates
 to true; otherwise, $p$ evaluates to false.  Using rules of logic we evaluate
-$r$, if $r$ true, then we add the course associated with it to $L$.  Once this
-is complete, $L$ will consist of every course that could be taken based solely
-on prerequisites that we have fulfilled.  \item We now prune $L$ by removing all
-courses that are not offered in the current quarter. $L$ now only contain
-classes that we can take given the current quarter and the classes we have
+$r$. If $r$ is true, we add the course associated with it to $L$.  Once this
+is complete, $L$ will consist of every course a student is eligible to take, based solely
+on satisfied prerequisites.  \item We now prune $L$ by removing all
+courses that are not offered in the current quarter. $L$ now contains
+only classes that we can take given the current quarter and the classes we have
 already taken.  \item We now construct a conflict graph, $G_{conflict}$, where
@@
 $i$ and $j$ if and only if the time of course offering $i$ overlaps the time of
-course offering $j$. An example conflict graph for one time period is given in
-Fig. \ref{conflictg}. \end{enumerate} 
+course offering $j$. Figure~\ref{conflictg} shows an example conflict graph for
+one time period. \end{enumerate}
 \begin{figure} [ht] 
@@
 \subsubsection{Analysis} The selection procedure involves a lot of set queries
-and iteration, but in asymptotic running time the algorithm is relatively quick.
-Let $n$ be the total number of classes ever available to take, or in other
-words, $n = \|U\|$. In the worst case, $P$ has a complex rule for all $n$
+and iteration, but in asymptotic running time the algorithm is efficient.
+Let $n$ be the total number of classes ever available to take.
+In the worst case, $P$ has a complex rule for all $n$
 courses, and checking each rule in $P$ is linear in the number of propositions.
@@
 rule has twelve propositions, so we can bound the rule check time by a constant,
-$C$, and therefore step one can be done in $O(n)$ time.
+$C$, and therefore step one runs in $O(n)$ time.
 
 Continuing to step two, the worst case is when $L$ = {all classes offered during
-the quarter}.  Checking whether a course is offered in a given quarter can be
-done in constant time, and we have $n$ courses to check, therefore step two can
-also be done in $O(n)$ time.
+the quarter}.  Checking whether a course is offered in a given quarter takes
+constant time, and we have $n$ courses to check, therefore step two also
+runs in $O(n)$ time.
 
@@
 \subsection{Selection of Quarterly Classes} To select classes we give each class
-a value and weight. The primary problem is the the classes contained in $L$
+a value and weight. The primary problem is the classes contained in $L$
 almost certainly have time conflicts, likely between courses that need to be
@@
 
-\subsubsection{Heuristic Value Function} In order to determine which classes
-should be picked in a given schedule, we use a heuristic function which assigns
-each class a value. The higher the value of this function for a class, the more
-likely that class will be placed in the schedule.  Recalling that our goal is to
+\subsubsection{Heuristic Value Function} To pick the classes taken in a given quarter,
+we use a heuristic function which assigns
+each class a value. When constructing the schedule, we pick classes with the highest heuristic value.
+Recall that our goal is to
 minimize the total number of quarters taken (and implicitly, maximize the total
-number of courses taken per quarter).  Our heuristic function should give higher
-value to class that maximize the number of classes taken this quarter and in
+number of courses taken per quarter). Therefore, our heuristic function gives higher
+value to class that maximize the number of classes taken in a given quarter and in
 future quarters.  One component to the heuristic function then must be the
 number of dependent courses a given course offering has. If a course satisfies
-many prerequisites, then we know that there are many courses that we cannot
-complete without it, so we want our heuristic function to favor it.
+prerequisites, then there are courses that we cannot
+complete without it. Therefore, we want our heuristic function to favor it.
 
@@
 $\beta$, and $\gamma$ are real numbers, giving a weight to piece of the
-heuristic. The scalars are determined experimentally.
+heuristic. We determine these scalars experimentally.
 
 \subsubsection{Selection Procedure} At this point in the algorithm we know which
-courses are available to a student and their relative values. We present two
+classes are available to a student and the values of the classes. We present two
 ways of approaching quarterly class choice.  We will initially encode the class
@@
 presented by Dinkel et al \cite{dinkel:scheduling}. and Pritsker et al
-\cite{prisker:informs}. A second approach developed by Yamada et al
-\cite{yamada:heuristic} and Pferschy et al \cite{pferschy:kcg} is presented for
-dealing with very large class choices where a binary LP would be difficult
-solve.
+\cite{prisker:informs}. We then present a second approach developed by Yamada et al
+\cite{yamada:heuristic} and Pferschy et al \cite{pferschy:kcg} for
+dealing with large class choices where a binary linear program is difficult to solve.
 
@@
 
-The linear program can be interpreted as follows: \begin{itemize} \item
+We interpret the linear program as follows: \begin{itemize} \item
 Eqn.~\ref{qtrval}: The value of a quarter is the total number of courses taken
@@
 represents a short hand way of indicating the number of credits hours a class
-takes. \item Eqn.~\ref{timecon}: Only one course can be taken in one particular
-time slot, this is where $G_{conflict}$ is used. Without the prebuilt conflict
-graph it would be difficult to build the set of constraints based on equation
-number~ref{timecon}. \item Eqn.~\ref{coursecon}: Previously in
-section~\ref{classDefs} we noted that in many cases a course is offered in
-multiple sections during the same quarter. A student can only take one offering
-of a course, this constraint eqn must be built for every class-quiz section
-combination which represent that same course.\end{itemize}
+takes. \item Eqn.~\ref{timecon}: A student can take only one class in one particular
+time slot. Recall that we use $G_{conflict}$ to determine time conflicts.
+\item Eqn.~\ref{coursecon}: Previously in
+section~\ref{classDefs} we noted that it is possible for a course to have multiple
+offerings during the same quarter. A student can only take one offering
+of a course during a quarter. We build this constraint for every class-quiz section
+combination for a given course.\end{itemize}
 
@@
 solve using modern LP solving software \cite{yamada:heuristic}. In our case,
-just solving for the optimal classes for a single major fits well into the scope
-of a few hundred (or even just tens of) variables. For multiple majors or to
+just solving for the optimal classes for a single major is on the order of
+hundreds of decision variables. For multiple majors or to
 consider class selection for distribution and elective requirements, however the
-number of classes balloons significantly and the KCG problem becomes much
-harder. For this version of the problem we suggest the techniques used by Yamada
+number of classes balloons and the KCG problem becomes infeasible.
+For this version of the problem we suggest the techniques used by Yamada
 et al \cite{yamada:heuristic}, which we briefly outline below.
@@
 In Yamada et al’s \cite{yamada:heuristic} numerical runtime analysis they found
-this algorithm ran very favorably against solving the same problem using popular
+this algorithm ran favorably against solving the same problem using popular
 linear programming solvers. There is one problem that we should briefly address,
-because the number of periods of available classes per week is relatively small
-the conflict rate for a large number of class is relatively higher than Yamada
+because the number of periods of available classes per week is small,
+the conflict rate for a large number of class is higher than Yamada
 et al’s 2\% maximal conflict rate for their experiments. Our data has
@@
 is between 1\% and 2\% of the figures used by Yamada \cite{yamada:heuristic} and
-also our total number of choices is in many cases (particular just the classes
+also our total number of choices is in some cases (particular just the classes
 in a particular major) less than 200. Giving the size and type of data in our

### 2010-12-13 paper/paper.tex
normalize sentence spacing; edit abstract
add determining dependents section, header, removed section 6

@@
 \renewcommand{\headrulewidth}{0pt} % customise the layout...  
-\lhead{}\chead{}\rhead{}
-\lfoot{}\cfoot{\thepage}\rfoot{}
+\lhead{Minimizing Time Required For Graduation}\chead{}\rhead{} \lfoot{}\cfoot{\thepage}\rfoot{}
 
@@
 that works towards completing a degree in the shortest amount of time 
-possible. The algorithm presented uses heuristics to attempt to pick the best
-class set each quarter, working towards a complete schedule quarterly by 
+possible.  The algorithm presented uses heuristics to pick the best
+class set each quarter, working towards a complete schedule quarter by
 quarter.}
@@
 To represent alternative and multiple prerequisites we introduced a set of logic
-rules for each course. The set $R$ consists of a prepositional logic statements representing
-the prerequisites of each course. The literals in each statement are a course.
-For example, the rule associated with $MATH327$ in Figure~\ref{logic_ex} is
-$$ \text{{\it MATH327: (MATH126 AND MATH300) OR MATH136}}.$$
+rules for each course.  The set $R$ consists of a prepositional logic statements
+representing the prerequisites of each course.  The literals in each statement
+are a course.  For example, the rule associated with $MATH327$ in
+Figure~\ref{logic_ex} is $$ \text{{\it MATH327: (MATH126 AND MATH300) OR
+MATH136}}.$$
 
@@
 we use a heuristic function which assigns
-each class a value. When constructing the schedule, we pick classes with the highest heuristic value.
-Recall that our goal is to
-minimize the total number of quarters taken (and implicitly, maximize the total
-number of courses taken per quarter). Therefore, our heuristic function gives higher
-value to class that maximize the number of classes taken in a given quarter and in
-future quarters.  One component to the heuristic function then must be the
-number of dependent courses a given course offering has. If a course satisfies
-prerequisites, then there are courses that we cannot
-complete without it. Therefore, we want our heuristic function to favor it.
+each class a value.  When constructing the schedule, we pick classes with the
+highest heuristic value.  Recall that our goal is to minimize the total number
+of quarters taken (and implicitly, maximize the total number of courses taken
+per quarter).  Therefore, our heuristic function gives higher value to classes
+that maximize both the number of classes taken in a given quarter and in future
+quarters.  One component to the heuristic function then must be the number of
+dependent courses a given course offering has.  If a course satisfies
+prerequisites, then there are courses that we cannot complete without it.
+Therefore, we want our heuristic function to favor it.
 
@@
 heuristic.  We determine these scalars experimentally.
+
+\subsubsection{Determining $Dependents(c)$} We briefly address how to determine
+the dependent value of a particular course. Figure~\ref{prereq} indicates the
+method for determining the dependent value of a course; generate the graph of
+prerequisites and then count the number of children in the subtree rooted at
+a particular course node. If this was the only consideration though, we could
+potentially assign a very high value to a course that is not related to
+a student's major becuase that course had a large number of dependents. The
+purpose of our system is to provide schedules that get a student through a major
+as quickly as possible, so we should only value courses that lead to graduation.
+The set $S$ describes the classes a student will need to graduate.
+
+To assign values to courses, we start with the courses in $S$ and work backwards
+using the prerequisite rules in $R$ to determine how to traverse the
+prerequisite DAG. Traversing the graph using this method, we determine the value
+of the course by counting the children in trees rooted at the current node in
+the traversal. By starting with the courses in $S$ we ensure that we only given
+values to courses that we need to take, instead of any courses with a large
+number of dependents.
 
@@
 
-\section{Computational Results} In this section we will outline results obtained
-by running computer simulations of schedule picking. We will computationally
-examine what parameters in our heuristic function produce the most desirable
-schedules.
-
 \bibliography{paper_references}{} \bibliographystyle{plain} \end{document}

### 2010-12-15 paper/paper.tex
Final revisions, more spelling and typograpical fixes, new assumptions section

@@
 prerequisites using a directed acyclic graph.  Each node represented a course, and edges
-existed between nodes $i$ and $j$ if and only if $i$ was a prerequisite for $j$.
+exist between nodes $i$ and $j$ if and only if $i$ was a prerequisite for $j$.
 This model works well if each course has only one prerequisite.  However,
@@
 
-Instead of trying to solve the complete problem we chose to use the "rolling
+Instead of trying to solve the complete problem we chose to use the ``rolling
 horizon approach" used by Xiangtong et al \cite{xiangton:informs} when
@@
 to the smaller time period, they increased the size of the time period, or rolled
-the horizon, and used the previously optimal solution to help determine
+the horizon, and used the previous optimal solution to help determine
 a solution to the new larger problem.  The team's algorithm continued rolling the
@@
 
+\subsection{Assumptions} Our method and in particular, our heuristic value
+function assumes continuity over time in how courses are offered. Since we do
+not have perfect schedule information, we don't look into the future, however we
+do assume that future schedules will bare resemblance to past schedules. The
+closer past course offering frequency and times offered is to the current
+schedule the better our predictions will be. We also assume that a student does
+not fail courses and therefore never needs to retake a course, that is once
+a course is entered in the set $T$ it never leaves.
+
 \subsection{Determining Available Course Offerings} We need to select the
@@
 is complete, $L$ will consist of every course a student is eligible to take, based solely
-on satisfied prerequisites.  \item We now prune $L$ by removing all
-courses that are not offered in the current quarter.  $L$ now contains
-only classes that we can take given the current quarter and the classes we have
-already taken.  \item We now construct a conflict graph, $G_{conflict}$, where
-every course offering in $L$ is a node, and there is an edge between two nodes
-$i$ and $j$ if and only if the time of course offering $i$ overlaps the time of
-course offering $j$.  Figure~\ref{conflictg} shows an example conflict graph for
-one time period.  \end{enumerate}
-\begin{figure} [ht] 
-    \begin{center}
-        \includegraphics[scale=0.35]{conflict_graph_ex} 
-    \end{center} 
-    \caption{A Conflict Graph} 
-    \label{conflictg} 
-\end{figure}
+on satisfied prerequisites.  \item We now prune $L$ by removing all courses that
+are not offered in the current quarter, and all courses which the student has
+already taken. Firs we take the set difference $L-T$ and the set difference
+between $L$ again and the input set of courses offered this quarter.  $L$ now
+contains only classes that we can take given the current quarter and the classes
+we have already taken.  \item We now construct a conflict graph, $G_{conflict}$,
+where every course offering in $L$ is a node, and there is an edge between two
+nodes $i$ and $j$ if and only if the time of course offering $i$ overlaps the
+time of course offering $j$.  Figure~\ref{conflictg} shows an example conflict
+graph for one time period.  \end{enumerate} \begin{figure} [ht] \begin{center}
+\includegraphics[scale=0.35]{conflict_graph_ex} \end{center} \caption{A Conflict
+Graph} \label{conflictg} \end{figure}
 
@@
 Let $n$ be the total number of classes ever available to take.
-In the worst case, $P$ has a complex rule for all $n$
-courses, and checking each rule in $P$ is linear in the number of propositions.
+In the worst case, $R$ has a complex rule for all $n$
+courses, and checking each rule in $R$ is linear in the number of propositions.
 However, we know {\it a priori} from analysis of our data set that the largest
@@
 function must therefore consider how often and how far in the future a given
-course is offered.  For example, if we are deciding between two course offering
+course is offered.  For example, if we are deciding between two classes
 that are identical except one is offered once a year, and the other is offered
@@
 
-We these considerations in mind, we begin to develop our heuristic function.  In
+In
 order to define our heuristic function, we must first define some auxiliary
-functions on course offering.  \begin{itemize} \item Let $TimeFromNow(c, q)$ be
-difference in quarters between the current quarter, $q$, and the next quarter
-a given course offering, $c$, is offered.  As there are four quarters, the range
+functions on course offerings.  \begin{itemize} \item Let $TimeFromNow(c, q)$ be
+difference in quarters between the current quarter, $q$, and the \emph{last} quarter
+a given course, $c$, was offered.  As there are four quarters, the range
 of this function is $\{0, 1, 2, 3\}$.  \item Let $Quarters(c)$ be the number of
 quarters per year a given course, $c$, is offered.  As there are four quarters,
-the range of this function is $\{0, 1, 2, 3, 4\}$.  \item Let $Dependents(c)$ be
+the range of this function is $\{0, 1, 2, 3\}$.  \item Let $Dependents(c)$ be
 the number of course that either directly or indirectly require a given course,
@@
 $\beta$, and $\gamma$ are real numbers, giving a weight to piece of the
-heuristic.  We determine these scalars experimentally.
+heuristic. Determining the best weights should be done experimentally.
 
@@
 potentially assign a very high value to a course that is not related to
-a student's major becuase that course had a large number of dependents. The
+a student's major because that course had a large number of dependents. The
 purpose of our system is to provide schedules that get a student through a major
@@
 prerequisite DAG. Traversing the graph using this method, we determine the value
-of the course by counting the children in trees rooted at the current node in
+of the course by counting the children in the tree rooted at the current node in
 the traversal. By starting with the courses in $S$ we ensure that we only given
@@
 \begin{equation}
-   X_i + X_j \leq 1, \forall \text{ pairs } (X_i,X_j) \text{ where } X_i \text{and} X_j \text{ are offerings of
+   X_i + X_j \leq 1, \forall \text{ pairs } (X_i,X_j) \text{ where } X_i \text{ and } X_j \text{ are offerings of
    the same course}
@@
 takes.  \item Eqn.~\ref{timecon}: A student can take only one class in one particular
-time slot.  Recall that we use $G_{conflict}$ to determine time conflicts.
+time slot, $t$  Recall that we use $G_{conflict}$ to determine time conflicts.
 \item Eqn.~\ref{coursecon}: Previously in
